### 2012-10-26 doc/rotation/abstract.tex
Added 'rotation' project document

\begin{abstract}
This is the abstract for this article.
\end{abstract}