### 2012-11-17 doc/rotation/abstract.tex
Added a rough introduction and fixed some typos

\begin{abstract}

There has been a series of results recently on recovering hidden parameters
(inference) from probabilistic models using the method of moments. While this
problem is in general hard, use of certain spectral techniques like
eigen-decomposition have resulted in tractable inference using only the first
few moments (3-4) of the data. This article will survey these results, attempt
to describe the approach in general and discuss possible issues with their
generality.

\end{abstract}