### 2013-01-17 doc/icml2013/abstract.tex
Added ICML 2013 draft

\begin{abstract} 
  Spectral algorithms for latent variable models have seen considerable
  recent interest for being efficient consistent estimators of model
  parameters. These algorithms make few if any assumptions about the
  generative process of the data, while providing a polynomial sample
  and computational complexity. We present a new spectral algorithm for
  a discriminative model, a mixture of linear regressions and show that
  it can recover the regression coefficients with similar polynomial
  guarantees. We evaluate the algorithm on linearly and non-linearly
  generated data, as well as on a motion tracking task and compare it's
  characteristics with an E-M algorithm.
\end{abstract}