### 2013-01-17 doc/icml2013/discussion.tex

\section{Discussion}
\label{sec:discussion}

Role of spectral algorithms in discriminative tasks.

Challenges and ways forward?