### 2013-02-14 doc/icml2013/abstract.tex
updates

\begin{abstract}

Discriminative latent-variable models are typically learned using
EM or gradient-based optimization, which suffer from local optima.
In this paper, we develop a new computationally efficient
and provably consistent estimator for the mixture of linear regressions,
a simple instance of discriminative latent-variable models.
Our approach relies on a low-rank linear regression to recover
a symmetric tensor, which can be factorized into the parameters
using the tensor power method.
We prove rates of convergence for our estimator
and provide an empirical evaluation illustrating
its strengths relative to local optimization (EM).

%  Spectral algorithms for latent variable models have seen considerable
%  recent interest for being efficient consistent estimators of model
%  parameters. These algorithms make few if any assumptions about the
%  generative process of the data, while providing a polynomial sample
%  and computational complexity. We present a new spectral algorithm for
%  a discriminative model, a mixture of linear regressions and show that
%  it can recover the regression coefficients with similar polynomial
%  guarantees. We evaluate the algorithm on linearly and non-linearly
%  generated data, as well as on a motion tracking task and compare it's
%  characteristics with an E-M algorithm.
\end{abstract}