### 2013-02-15 doc/icml2013/discussion.tex

\section{Conclusion}
\label{sec:conclusion}

Role of spectral algorithms in discriminative tasks.

Challenges and ways forward?