### 2013-02-15 doc/icml2013/discussion.tex
conclusion

\section{Conclusion}
\label{sec:conclusion}

In this paper, we take a first step at developing computationally efficient and
statistically consistent estimators for discriminative latent-variable models,
by focusing on mixture of linear regressions.  We showed that by regressing on
higher order powers of the data with a regularizer that encourages low rank
structure, coupled with an efficient tensor power method, we obtain a very natural algorithm.
Empirically, our algorithm can outperform EM, and when used to initialize EM,
yields significantly more accurate and stable results.