### 2013-02-16 doc/icml2013/discussion.tex
edit

\section{Conclusion}
\label{sec:conclusion}

In this paper, we developed a computationally efficient and statistically
consistent estimator for mixture of linear regressions, a first step at
tackling discriminative latent-variable models.  Our algorithm, Specral
Experts, regresses on higher-order powers of the data with a regularizer that
encourages low rank structure, followed by tensor factorization to recover the
actual parameters.  Empirically, Spectral Experts generally does not outperform
EM, but when used to initialize EM, yields significantly more accurate and
stable results than EM.