### 2013-04-01 doc/icml2013/rebuttal.txt
Inclusions for the rebuttal round

Thank you for your constructive comments and suggestions; we will try to
respond to your questions below.

# Major Comments:

## Reviewer 30 (Regarding Lemma 3):
The $O(1/\sqrt{n})$ convergence follows from applying Hoeffding's
inequality on each term of the random vectors $\eta_p(x) x$, which have
zero mean when taking expectation over $h$ and $\epsilon$. For example,
in $\eta_2(x) x$, we have three terms (Equation 3), which each have
expectation zero by independence of $x$, $h$ and $\epsilon$.

This does not require that the $x$ have zero mean, and note that we do
not take expectations over $x$ at any point in our analysis.

## Reviewer 4 (Regarding analysis with whitening):
Your point is well noted; results from standard techniques (Hsu
& Kakade, 2013) show that the whitening step introduces an additional
factor of $O(\sigma_{min}(M_2)^{-1.5} \|M_2\|_{op}^{0.5} \|M_3\|_{op})$.
We will include this as a lemma in the final version as we can not
upload an updated version in the rebuttal phase.

## Reviewer 30 (Regarding convergence rate in experiments):

One of the obstacles in applying the bounds we derived is that it
requires the error of the empirical moment estimates to be less than
$\sigma_{\min}(M_2)/2$, a condition that did not hold true for our
experiments. For the data in Figure 3, this condition required the
moment errors to be less than 0.15, while the error in estimates we
recovered were at least 0.9. We have also identified that one major
reason for this was that the proximal gradient algorithm we used did not
converge in the specified number of iterations (i.e. 500).

# Minor Comments:

## Reviewer 30 (Regarding Identifiability):
Identifiability is only a problem when the input features are not
algebraically independent, i.e. each coordinate can not be expressed as
a deterministic analytic function of the remaining coordinates. Thus, it
is identifiable when the coordinates are generated from a random
Gaussian (or any other distribution).

## Reviewer 4 (Regarding Lemma 3):
We do not need to know $h$; in fact we are trying to bound the
difference between mean, $M_1$ which we recover by regression, and the
actual value of $\beta_h$ at each data point; it is useful to think of
it as bounding the variance of the zero-mean term $M_1 - \beta_h$.

## Reviewer 30 (Regarding operator norm for vectors):
This was a typo, and we meant the 2-norm.

## Reviewer 4 (Regarding references):
Thank you, we will include them in the final version.

## Reviewer 4 (Regarding $\sigma$):
We used $\sigma = 0.1$ for the results we presented, but found that they
were not very sensitive to the noise level.

## Reviewer 4 (Regarding Regularization):
We set the regularization parameters by grid search for $n = 10^5$ and
used the $1/\sqrt{n}$ rate to extrapolate for other $n$.

## Reviewer 7 (Regarding Generality of Experiments):
We found that robust initialization happened for almost all inputs. The
cases where we did not get robust initialization occurred when we did
not have good enough moment estimates and thus poor parameter estimates.
Furthermore, this problem went away when we used more samples.

## Reviewer 7 (Regarding Plot of Convergence):
We have compared the convergence of spectral experts and EM to the exact
parameters for increasing values of $n$ in Figure 3 in terms of the parameter
error.