### 2013-05-09 doc/icml2013/discussion.tex
finish experiments

\section{Conclusion}
\label{sec:conclusion}

In this paper, we developed a computationally efficient and statistically
consistent estimator for mixture of linear regressions.
Our algorithm, Spectral
Experts, regresses on higher-order powers of the data with a regularizer that
encourages low rank structure, followed by tensor factorization to recover the
actual parameters.  Empirically, we found Spectral Experts 
to be an excellent initializer for EM.
%yielding significantly more accurate and stable results than EM.