### 2013-05-21 doc/nips2013/abstract.tex
first draft of NIPS paper

\begin{abstract}
Latent-variable models.
\end{abstract}