### 2013-05-22 doc/nips2013/discussion.tex

\section{Related work}

% Observable operator models
HMMs \cite{hsu09spectral},
low-treewidth graphs \cite{parikh12spectral},
probabilistic context-free grammars \cite{cohen12pcfg},
weighted finite state automata \cite{balle12automata},
finite state transducers \cite{balle11transducer}.

% Method of moments
Latent Dirichlet Allocation \cite{anandkumar12lda},
mixture of spherical Gaussians \cite{hsu13spherical},
latent parse tree models \cite{hsu12identifiability}
mixture of linear regressions \cite{chaganty13regression}.
linear Bayesian networks \cite{anandkumar12linear}.

% Structure recovery
structure recovery
\cite{anandkumar11tree}

% Alternate estimators
There are multiple consistent estimators.
Composite likelihood \cite{besag75pseudo,lindsay88composite,liang08asymptotics}
We are assuming inference is tractable.
Our main problem is non-convexity.

\section{Conclusion}

That's it.