### 2013-05-22 doc/nips2013/setup.tex

\section{Setup} \label{sec:setup}

Let $z$ be a collection of variables indexed by $V$ (that is, $z = \{z_j\}_{j \in V}$).
Suppose that some of the variables are observed ($x = \{ z_j \}_{j \in O}$, where $O \subset V$) 
and the rest are latent ($h = \{ z_j \}_{j \in H}$, where $H = V \backslash O$).
Let $C \subset 2^V$ be a collection of subset of variables,
and let $\phi_c(z_c) \in \R^d$ be the feature map for each $c \in C$.
In this paper, we will consider both directed and undirected models over $z$.
In the undirected case, we define the standard globally-normalized exponential family model:
\begin{align}
  p_\theta(z) = \exp\{ \phi(z)^\top\theta - A(\theta) \}, \aside{undirected}
\end{align}
where $\phi(z) \eqdef \sum_{c \in C} \phi_c(z_c) \in \R^d$ is the global feature vector, $\theta \in \R^d$ is the parameter vector,
and $A(\theta) = \log \int \exp\{\phi(z)^\top\theta\} dz$ is the global log-partition function.

In the directed case, we assume a directed acyclic graph over $V$,
and define $\pi(j) \subset V$ to be the parents of each node $j \in V$.
Next, define a model as a product of locally-normalized models:
\begin{align}
  p_\theta(z) = \prod_{j \in V} p_\theta(z_j \mid z_{\pi(j)}), \quad\quad
  p_\theta(z_j \mid z_{\pi(j)}) = \exp \{ \phi_j(z_j, z_{\pi(j)})^\top\theta - A_j(\theta; z_{\pi(j)}) \}.
\end{align}
Here, each local model has its own local features $\phi_j(z_j, z_{\pi(j)})$
and its own local log-partition function $A(\theta; z_{\pi(j)})$, which depends on values of its parent.
To make the directed and undirected models more comparable,
define $\phi(z) = \sum_{j \in V} \phi(z_j, z_{\pi(j)})$
and $A(\theta; z) = \sum_{j \in V} A_j(\theta; z_{\pi(j)})$,
so that:
\begin{align}
  p_\theta(z) \eqdef \exp\{ \phi(z)^\top\theta - A(\theta; z) \}. \aside{directed}
\end{align}

%To some extent, our methods only depend on the conditional independence
%structure of the models, not on whether the model is directed or undirected.
%Later, we will show how to estimate

%\paragraph{Example.}
%For example, \reffig{simpleModels} Markov random fields are an instance.

\paragraph{Problem statement.}

% Statement
This paper focuses on the problem of parameter estimation:
We are given $n$ i.i.d.~examples of the observed variables $D = (x^{(1)}, \dots, x^{(n)})$
where each $x^{(i)} \sim p_{\theta^*}$ for some true parameters $\theta^*$.
Our goal is to produce a parameter estimate $\hat\theta$ which is hopefully
close to $\theta^*$.

% Maximum likelihood
The standard estimation procedure is maximum (marginal) likelihood,
$\hatthetaml \eqdef \max_{\theta \in \R^d} \sum_{x \in D} \log p_\theta(x)$,
which is statistically efficient but computationally intractable.
In practice, one uses EM or gradient-based optimization procedure on the marginal likelihood,
which can get stuck in local optima.

%Since both models have the same sufficient statistics lead to convex
%likelihoods (assuming all variables are observed), in many cases our methods
%will not be particular to which.  In some cases, however (e.g., the factorial models),
%we will rely on marginalization properties of the directed models.

% Principles
%By insisting on developing a consistent estimator and considering identifiability,
%The key is that a variable in the model must correspond to a bottleneck.