### 2013-05-31 doc/nips2013/factorial.tex

\section{Unshuffling factorization for factorial mixture models} \label{sec:factorialModels}

So far, we have relied on the existence of a bottleneck hidden variable $h_j$ on which
we have three conditionally independent views.  Such bottlenecks do not always
exist; for example, in \reffig{factorialModels}(a), neither $h_1$ or $h_2$ is a bottleneck.
Our solution is to create a composite latent variable $c$ which is in
one-to-one correspondence with the existing latent variables $(h_1,h_2)$.
We can invoke $\TensorFactorize$ to obtain composite marginals and
conditional means.  However, the identity of $h_1$ and $h_2$ becomes lost in the composite,
and must be recovered.  To do this, we develop a new algorithm $\UnshuffleFactorize$
to produce constraints on the original parameters.

\Fig{figures/factorialModels}{0.3}{factorialModels}{The factorial mixture model (a)
does not have three-view bottlenecks, so we first identify composite bottlenecks,
and then perform an \emph{unshuffling factorization} (b):
Given known composite conditional means $B_c = \BP[x_j \mid c] \in \R^\ell$,
which are derived from combinations of unknown parameters $f_a = \theta_1(\cdot, a), g_b = \theta_2(\cdot, b)$,
we group differences into bins (dashed boxes), and maximal bins into sources (red rounded boxes),
producing a linear program with variables $f_a,g_b,Z_{ab}$ and constraints for each $B_c$.
}

%restricted Boltzmann machines \cite{salakhutdinov09softmax}

We focus on the factorial mixture model.
Let $h = (h_1, \dots, h_s) \in [\nh]^s$ be the latent variables (sources)
and $x = (x_1, x_2, x_3) \in [\nphix]^3$ be the observed variables.
\begin{align}
p_\theta(x, h) = \prod_{i=1}^s p_\theta(h_i) \prod_{j=1}^3 p_\theta(x_j \mid h), \quad
p_\theta(x_j \mid h) = \exp\left\{ \sum_{i=1}^s \theta_i(x_j, h_i) - A(\theta; h) \right\},
\end{align}
where each $\theta_i \in \R^{\nphix \times \nh}$ are the parameters for the $i$-th source.
%Let $\theta = (\theta_1, \dots, \theta_b) \in \R^{\nphix \times b \nh}$.

Let $B \eqdef (\BP[x_j \mid c])_{x_j \in [\nphix],c \in [\nh]^s} \in
\R^{\nphix \times \nh^s}$ be the conditional mean of $x_j$ given composite cluster $c$.

\begin{assumption}
  Assume that $B$ has full column rank ($k^s$).\footnote{Note that this implies
  $d \ge k^s$, which is a rather stringent assumption, but is necessary for the
  three-view method of moments technique to work.}
\end{assumption}

First, we run $\TensorFactorize$ on estimates of
$M_2 \eqdef \E[x_1 \otimes x_2]$ and $M_3 \eqdef \E[x_1 \otimes x_2 \otimes x_3]$,
which produces $B = (\E[x_j \mid c])_{c \in [\nh]^s}$.
%Next, we still have to extract the actual parameters $\theta$ from $B$. 
In the simple mixture model,
the clusters produced by $\TensorFactorize$ are in exact one-to-one correspondence
with the actual mixture components, but in the factorial case, we must solve
a \emph{credit assignment problem} to match combinations of $s$ sources
with columns of $B$.

Let $\{ v_{ia} \}_{i \in [s], a \in [k]}$, where $v_{ia} = \theta_i(\cdot, a) \in \R^\ell$,
be the collection of parameter vectors that we seek to recover.
Let $C \in \R^{\ell \times k^s}$ be the elementwise logarithm of $B$; that is
$C(x_j, h) = \log B(x_j, h)$.
The columns of $C$ are exactly the unordered collection of vectors
$\{ \sum_{i=1}^s v_{ia_i} - Z_{\ba} : \ba = (a_1, \dots, a_s) \in [\nh]^s \}$,
where $Z_{\ba} = \log \sum_{e=1}^\ell (\sum_{i=1}^s \exp(v_{ia_i}(e)))$ is the normalization constant (a scalar).

Note that there is a combinatorial dimension to this problem since the association between
columns of $C$ and combinations of $\{ v_{ia} \}$ is unknown.
As a running example (\reffig{factorialModels}(b)),
consider $s=2$ sources, each taking on $k=3$ values (let $f_a = v_{1a}$ and $g_b = v_{2b}$).
The following algorithm creates a linear program which puts constraints on variables $\{v_{ia}\}$ and $Z_{\ba}$:
\begin{enumerate}

\item Compute the difference vector $B_i - B_j$ for every distinct pair of columns $c,d \in [\nh^s]$.

\item Put the pairs $(c,d)$ into bins, where the difference vectors $B_c-B_d$
in each bin are \emph{comparable}, by which we mean that they
differ by a constant scalar times $\bone$ (for the normalization constant).

\item Keep only the maximal bins.
These bins should contain $\nh^{s-1}$ pairs, each corresponding to a fixed setting of the $s-1$ sources
and changing the value of one source.
Specifically, the difference vector is equal to $v_{ia} - v_{ia'}$ up to a
constant scalar for some source $i \in [k]$ and distinct values $a, a' \in
[\nh]$.

\item Group the bins into sources:
  For each source $i = 1, \dots, s$,
  chose any bin, declare it as $v_{i1} - v_{i2}$, and add it to the source.
  Let $\gamma$ be the set of $c$ such that that bin contains some difference vector $B_c - B_d$,
  and include any other bin with the same $\gamma$; these should correspond to $v_{i1} - v_{ia}$ for $a \ge 3$.
  Then throw away any bin whose difference vector is $-(B_c - B_d)$
  or is comparable to the difference of two existing difference vectors (e.g., $v_{i2} - v_{i1}$ or $v_{i2} - v_{i3}$).
  The resulting source should have $\nh-1$ bins, corresponding to $\{ v_{i1} - v_{ia} : 2 \le a \le \nh \}$.
  At this point, we have identified which vectors contribute to each $B_c$,
  taking care of the combinatorial aspect of the problem,
  but have yet to recover the actual vectors.

  \item To do this last step, construct a linear program with $\nh^s \cdot d$ constraints, one for each entry of $B$.
  It will have $s\nh \cdot d + \nh^s$ variables,
  one for each of $s\nh$ vectors $v_{ia}$ and $\nh^s$ for the normalization constants $Z_\ba$.
  Note that $Z_\ba$ is actually a determinstic function of the the vectors, but since it is a non-linear one,
  it is difficult to work with; therefore, we create a separate variable to keep the problem linear.

  It turns out that this is LP is rank deficient for two reasons, the first intrinsic to our model,
  and the second due to loose linearization.
  First, the vectors $v_{ia}$ are non-identifiable since we can always add any vector $u \in \R^\ell$ to
  all the vectors of a source $i_1$ ($v_{i_1 a}$ for all $a \in [\nh]$)
  subtract $u$ from all the vectors of another source $v_{i_2 a}$ for all $a \in [\nh]$.
  Second, if we knew the normalization constants $\{Z_\ba\}$, then this is only rank deficiency.
  However, treating $\{Z_\ba\}$ as independent variables creates additional rank deficiency.

  Despite this, in the spirit of providing partial information,
  $\UnshuffleFactorize$ still gives us linear constraints on the parameters.
  In particular, we propose adding to the likelihood objective a strong regularization term
  $R(\theta, Z) = \frac{\lambda}{2} (\bA_1 \theta - \bA_2 Z - B)^2$, where $\bA_1$ and $\bA_2$ are coefficients
  recovered by unshuffling.
  In particular, we propose optimizing $\min_{\theta, Z} \sum_{i=1}^n -\log p_\theta(x^{(i)}) + R(\theta, Z)$.
  Although the resulting objective is still non-convex, the regularization imposes quite strong constraints,
  and the optimization would happen effectively over a much lower-dimensional subspace.
\end{enumerate}