### 2013-05-31 doc/nips2013/abstract.tex
made a pass up through section 4 (measurements)

\begin{abstract}
Parameter estimation in log-linear latent-variable models is typically
performed using gradient-based optimization on the log-likelihood, which is prone to local optima.
In this paper,
%we show how to leverage recent advances in the method of moments
%to alleviate these problems.
we leverage recent advances in the method of moments to estimate latent
moments of the target distribution,
which we incorporate as constraints on the original optimization
problem.
In favorable cases, these constraints provide full information and render the
original optimization problem
convex; in other cases, the constraints offer only partial information,
but still appear to provide significant aid to the original optimization
problem.
Factorial models do require special attention;
to handle these models,
we introduce a new technique called unshuffling factorization to induce
linear constraints on the parameters.
Empirically, our approach provides improvements on several models.
\end{abstract}