### 2013-05-31 doc/nips2013/measurements.tex

\section{Measurements: partial information from moments} \label{sec:generalModels}

Our learning algorithm for three-view mixture models relies on the fact that
the sufficient statistics of the exponential family
can be expressed as a function of the latent moments.
Of course, these conditions do not hold for general log-linear models.
Rather than falling back on plain local optimization of the likelihood in these
cases, we would still like
to somehow use the \emph{partial information} offered by the latent moments.
%We will show how to incorporate these moments into a likelihood-based
%objective.
Partial information is available when the following holds:
\begin{property}
  There exists some latent variable $h_{j_0} \in [\nh]$ (the bottleneck)
  and three conditionally independent observed variables $x_{j_1},x_{j_2},x_{j_3}$ (views) such that
  each conditional mean matrix $\E[\phix(x_{j}) \mid h_{j_0}] \in \R^{\nphix \times \nh}$ ($j \in \{ j_1, j_2, j_3 \}$) has full column rank $\nh$
  for some transformation $\phix$.
\end{property}

% Examples
\reffig{generalModels} shows examples of log-linear models.
For notational simplicitly, assume each latent variable $h_j \in [\nh]$ is an integer
and each observed variable $x_j \in \R^\ell$ is an indicator vector encoding
one of $\ell$
possible values.

\Fig{figures/generalModels}{0.3}{generalModels}{
Three examples of undirected log-linear models.  In each of these cases,
we can recover latent moments, which are themselves not sufficient to identify the parameters,
but provide partial information which leads to a more well-behaved likelihood function.
}

\paragraph{(a) HMM}
Following the reduction in \citet{anandkumar12moments},
we could take $h_2$ as the bottleneck and $x_1, x_2, x_3$ as the three views.
Define $\pi = (\BP[h_1]) \in \R^{\nh}$,
transition matrix $T = (\BP[h_3 \mid h_2]) \in \R^{\nh \times \nh}$
and observation matrix $O = (\E[x_2 \mid h_2]) \in \R^{\nphix \times \nh}$.
$\TensorFactorize$ recovers the marginal probability of the bottleneck
$T \pi$ and conditional means of the three views
$O T^\top \diag(T \pi)^{-1}$,
$O$,
and $O T$.
Note that we can recover $T = O^\dagger(OT)$ and $\pi = T^{-1} (T\pi)$,
from which we can compute the latent moments/sufficient statistics $\mu = \E[\phi(x,h)]$.

However, this model only uses the first three observations,
which doesn't use all the information available.
Let us consider a new reduction:
Let $h_j$ be the bottleneck where $j$ is a \emph{random} variable
taking on values uniformly from $\{2,
\dots, L-1\}$, where $L$ is the number of time steps.
Let $x_{j-1},x_j,x_{j+1}$ be the three views.\footnote{
When we actually estimate the observed moments, we simply average over the choice of $j$.}
Now, note that we can recover $O = \inv{L-2} \sum_{j=2}^{L-1} \E[x_j \mid h_j]$ from the second view,
but using the data more fully.  The conditional means of the first and third
views are now a complex function of the parameters.
%due to marginalization over the rest of the variables,
However, the point is that we can get most of our utility out of the partial
information offered by $O$ and ignore information which is hard to ingest.

\paragraph{(b) Grid model}
Analogous to the second HMM reduction, let $h_{ij}$ be the bottleneck,
where $(i,j)$ is chosen randomly.
Let $x_{ij}^a,x_{ij}^b$ be two of the views, and any other observed variable as the third view.
$\TensorFactorize$ would reveal conditional means $\E[x_{ij}^a \mid h_{ij}]$
and $\E[x_{ij}^b \mid h_{ij}]$,
and a third quantity that involves the latent-to-latent features and the observation features.
This quantity involves marginalizing out the rest of the graph,
and therefore difficult to use.
%to extract something useful like $\BP[h_{ij}, h_{i(j+1)}]$
%in closed form.
%Note that the analog of this was still possible in the first HMM reduction that
%used only the first three observations, but not in the second one.

\paragraph{(c) Tall mixture model}
Let $h_0$ be the bottleneck and $x_1,x_2,x_3$ be the three
views.
$\TensorFactorize$ reveals the conditional means $\E[x_j \mid h_0]$, which are
not even expectations over features in the original model, let alone sufficient
statistics.\footnote{Note that
the model might not be identifiable in general \cite{anandkumar11tree}.
However, the model is identifiable with adequate parameter sharing,
and therefore
parameter estimation in this model is still meaningful.}

In each of these three models, we could extract marginals $\pi = \BP[h_j] \in \R^{\nh}$
and some conditional mean matrix $B = \E[x_{j'} \mid h_j] \in \R^{\nphix \times \nh}$,
but these are insufficient to identify the parameters.
How can we still use this partial information $(\pi,B)$?
To this end, we adopt the measurements framework of \citet{liang09measurements},
which allows us to incorporate moment constraints into a likelihood-based objective in a coherent way.
The basic idea is to define two families of distributions:
$\sQ$, which are consistent with all information; and
$\sP$, the exponential family.  The
framework tries to choose $q \in \sQ$ and $p \in \sP$ to minimize
$\KL{q}{p}$.

Specifically,
let $\bx = (x^{(1)}, \dots, x^{(n)})$ be the vector of observed variables
and let
$\bh = (h^{(1)}, \dots, h^{(n)})$ be the vector of latent variables.
Our model family $\sP$ is defined by product distributions
$p_\theta(\bx,\bh) = \prod_{i=1}^n p_\theta(x^{(i)}, h^{(i)})$ as $\theta \in \R^d$.
Suppose our observations are $(\tilde x^{(1)}, \dots, \tilde x^{(n)})$
and we have recovered $(\pi,B)$, where $\pi$ is the marginal of the bottleneck $h_j$ and $B$ is the conditional mean of $x_{j'}$.

Define $\sQ$ to be $q$ satisfying the following constraints:\footnote{We use a hard constraint
for expository clarity, but in practice, we use a quadratic penalty.  See
\citet{liang09measurements} for details.}
\begin{align}
  q(x^{(i)} = \tilde x^{(i)}) &= 1, \quad i \in [n] \label{eqn:exConstraint} \\
  \inv{n} \sum_{i \in [n]} q(x_{j'}^{(i)} = v, h_j^{(i)} = u) &= \pi_u B_{v u}, \quad u \in [\nh], v \in [\nphix]. \label{eqn:momentConstraint}
\end{align}
The first constraint $\sQ$ encodes our observed examples
and the second puts additional constraints on the latent variables.
Our final objective function is as follows:
\begin{align}
  \min_{q \in \sQ} \min_{\theta \in \R^d} \KL{q(\bx, \bh)}{p_\theta(\bx, \bh)}. \label{eqn:minKL}
\end{align}
\paragraph{Intuitions.}
With only the first constraint \refeqn{exConstraint}, the objective
\refeqn{minKL} is equivalent to our original maximum (empirical) marginal
likelihood problem, which is non-convex.  If the second constraint
\refeqn{momentConstraint} are exactly the features/sufficient statistics of the
model (as in the three-view mixture model from \refsec{threeViewMixtureModel}),
then \refeqn{minKL} is always convex and
converges to the (expected) maximum likelihood objective function as the number of examples $n \to\infty$.

% Optimization
We optimize \refeqn{minKL} using a variant of EM.
Briefly, the E-step is equivalent to a generalized maximum entropy problem,
which we solve by taking the dual, resulting in a parameter for each moment constraint \refeqn{momentConstraint}.
In the M-step, we maximize the expected log-likelihood with respect to $q$.

%$\beta = \argmax_{q} H(q) \tau^\top\beta - B(\beta, \theta) + r(\beta)$
%E-step: $\beta = \arg\max_{\beta} \tau^\top\beta - B(\beta, \theta) - r(\beta)$
%M-step: $\theta = \arg\max_\theta \E_q[\phi(\bx,\bh)^\top - A(\theta)]$.