### 2013-05-31 doc/nips2013/mixture.tex

\section{Three-view mixture models}
\label{sec:threeViewMixtureModel}

\Fig{figures/simpleModels}{0.3}{simpleModels}{Three-view mixture models: conditioned on a latent variable $h \in [\nh]$,
we have three conditionally independent observed variables.  We consider
estimation for both directed (left) and undirected (right) versions.
}

In this section, we consider the simplest model to illustrate the form of our
method.  Consider the undirected mixture model model in \reffig{simpleModels}(b).
We have one discrete latent variable $h \in [\nh]$ and
three observed variables
$x = (x_1, x_2, x_3) \in \sX_1 \times \sX_2 \times \sX_3 = \sX$
which are conditionally independent given $h$.
Formally, the model is as follows:\footnote{We assume for simplicity that the $x_j$'s are identically distributed conditioned on $h$,
but it is possible to use different distributions (see \citet{anandkumar12moments}).}
\begin{align}
  \label{eqn:mixtureModel}
  p_\theta(x, h) \eqdef \exp \{ \phi(x, h)^\top\theta - A(\theta) \}, \quad\text{where} \, \phi(x, h) \eqdef \phiprior(h) + \sum_{j=1}^3 \philocal(x_j, h) \in \R^d.
\end{align}

We will now describe the algorithm for producing a parameter
estimate $\hat\theta$.  For simplicity of exposition, we will assume we have an infinite
amount of data and work with population quantities.  Actual parameter
estimates are constructed by simply plugging in the empirical versions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Step 1: Compute observed moments (aggregation)}

%Rather than directly trying to connect the parameters $\theta$ to the data $D$,
Let us start bottom-up by simply computing properties of the data.
Let $\phix : \sX \mapsto \R^\ell$ denote a \emph{transformation function}
(the requirement on $t$ will be specified later),
which aims to extract properties of $x$ relevant to $h$.
Define the \emph{observed moments} (analogous to those used in \cite{anandkumar12moments}):
\begin{align}
  M_2 \eqdef \E[\phix(x_1) \otimes \phix(x_2)] \in \R^{\nphix \times \nphix}, \quad\quad
  M_3 \eqdef \E[\phix(x_1) \otimes \phix(x_2) \otimes \phix(x_3)] \in \R^{\nphix \times \nphix \times \nphix}.
\end{align}
Importantly, these quantities are easily estimated from data by taking empirical averages.

\subsection{Step 2: Recover latent moments (factorization)}

By exploiting the conditional independence structure of the model, the observed
moments $M_2,M_3$ (which only reference the observed variables) admit a
factorization in terms of the \emph{latent moments}
(which provides a window into the latent variables) \cite{anandkumar12moments}.

Define
$\pi \eqdef (\BP[h=1], \dots, \BP[h=\nh])$ to be the marginal probabilities
over $h$,
and define the matrix $B \eqdef (\E[\phix(x_j) \mid h])_{h \in [\nh]} \in \R^{\ell \times k}$,
where each column is the expected transformated data conditioned on $h$.

Now we can express the observed moments $(M_2,M_3)$ in terms of the latent moments $(\pi,B)$:
\begin{align}
  M_2 = \sum_h \pi_h B_h \otimes B_h \in \R^{\nphix \times \nphix}, \quad\quad
  M_3 = \sum_h \pi_h B_h \otimes B_h \otimes B_h \in \R^{\nphix \times \nphix \times \nphix}.
\end{align}
\begin{assumption}
  The latent moments matrix $B \in \R^{\nphix \times \nh}$ has full column rank $\nh$.
\end{assumption}

\citep{anandkumar13tensor} presented a robust tensor power method, which we call $\TensorFactorize$,
which converts consistent estimates of $(M_2,M_3)$ into consistent estimates of $(\pi, B)$ (up to permutation of the columns).
In brief, the algorithm first computes a whitening matrix $W \in \R^{\nphix \times \nh}$ such that $W^\top M_2 W = I_{k \times k}$,
and uses it to whiten $M_3$.  Then a robust tensor power method is applied to
extract the eigenvectors of the whitened $M_3$;
unwhitening yields the columns of $B$.
%\begin{lemma}[tensor factorization \citep{anandkumar13tensor}]
%Let $B \in \R^{\nphix \times \nh}$ be a rank $k$ matrix.
%Let $M_2 = \sum_h \pi_h B_h \otimes B_h \in \R^{\nphix \times \nphix}$ be a rank $k$ symmetric matrix.
%and $M_3 = \sum_h \pi_h B_h \otimes B_h \otimes B_h \in \R^{\nphix \times \nphix \times \nphix}$ be a rank $k$ symmetric tensor.
%There exists an polynomial time algorithm $\TensorFactorize$ that takes consistent estimates of $M_2$ and $M_3$
%and returns consistent estimates of $(\pi, B)$ (up to permutation of the columns).
%\end{lemma}

We need to specify conditions on the transformation function $\phix$
to ensure that enough information is retained to pinpoint the parameters later.
The following condition characterizes $\phix$:
\begin{property}
  \label{property:transformation}
  The transformation $\phix$ (chosen by us) is such that $C \eqdef
  (\E[\philocal(x_j, h) \mid h])_{h \in [\nh]} \in \R^{d \times h}$ is a
  function of $B$.
\end{property}
Here, $C$ provide a view on the local features $\philocal(x_j,h)$,
which are
closely related to the model parameters \refeqn{mixtureModel},
and $B$ is what we have recovered so far.
%Note that we couldn't recover $C$ directly because it depends on $h$.

It is easy to construct $\phix$ such that Property~\ref{property:transformation} holds.
Since $h$ is discrete,
we can always take $\phix(x) = \flatten((\philocal(x, h))_{h \in [\nh]}) \in \R^{dk}$,
which contains $k$ blocks, one for each latent state.
Then each column $C_h \in \R^d$ is simply the $h$-th block
of the column $B_h \in \R^{kd}$.
%We can recover a column $C_h = \E[\philocal(x_j, h) \mid h]$
%Each column $C_h = \E[\philocal(x_j, h) \mid h]$
%by taking the block of column $B_h = \E[\flatten(\philocal(x_j, h'))_{h' \in [\nh]} \mid h]$
%corresponding to $h = h'$.
More simply, if
the local features are simply the tensor product of some function $g(x)$ and an
indicator on $h$ ($\philocal(x,h) = \flatten(g(x) \otimes e_h)$),
then $\phix(x) = g(x)$ suffices.

\subsection{Step 3: Recover parameters (optimization)}

In \cite{anandkumar12moments}, our latent moments $(\pi, B)$
would be the parameters, but for log-linear models,
we still need to do a bit more work.
First, note that we can express the sufficient statistics $\E[\phi(x,h)]$
of the model in terms of the quantities recovered in Step 2:
\begin{align}
  \mu \eqdef \E[\phi(x,h)] = \E[\phiprior(h)] + \sum_{j=1}^3 \E[\philocal(x_j,h)] = \sum_{h \in [\nh]} \pi_h (\phiprior(h) + 3 C_h).
\end{align}
By properties of the exponential family, $\mu$ suffices to recover $\theta$.
To elaborate, if we had infinite \emph{labeled} examples $(x,h)$,
then we would seek to maximize the joint log-likelihood:
\begin{align}
L(\theta) \eqdef \E[\log p_\theta(x, h \mid \theta)] = {\underbrace{\E[\phi(x,h)]}_{\mu}}^\top\theta - A(\theta).
\end{align}
Note that the optimization problem here only depends on the data via $\mu$.
Instead of estimating $\mu$ from labeled examples,
we estimate $\mu$ from tensor factorization applied to unlabeled examples.

We thus obtain a consistent estimator for $\theta$ assuming the exponential
family is identifiable from fully-observed data (that is
the log-artition function $A(\theta)$ is strongly convex).
%\begin{assumption}
%The log-linear model is identifiable from fully-observed data.
%Equivalently, the log-partition function $A(\theta)$ is strongly convex.
%\end{assumption}
%Identifiability means that the moments $\mu$ uniquely determine $\theta^*$.
%By the moment-generating properties of the log-partition function,
%the optimality condition is that $\mu = \nabla A(\theta) = \E_\theta[\phi(x, h)]$.
In other words,
tensor factorization (Step 2) yields the mean parameters of the exponential family
($\mu = \nabla A(\theta) = \E_\theta[\phi(x, h)]$),
which can be mapped to (canonical) parameters $\theta$ via convex optimization.

% Directed mixture model
\paragraph{Directed mixture model}

The directed mixture model, we have $p_\theta(x,h) = p_\theta(h) \prod_{j=1}^3 p_\theta(x_j \mid h)$,
where
$p_\theta(h) \eqdef \exp\{\phiprior(h)^\top\theta - \Aprior(\theta)\}$
and
$p_\theta(x_j \mid h) \eqdef \exp\{\philocal(x_j, h)^\top\theta - \Alocal(\theta; h)\}$.
Consequently, the joint log-likelihood is:
\begin{align}
  L(\theta) = {\underbrace{\E[\phi(x,h)]}_{\mu}}^\top\theta - \E[A(\theta; z)], \quad
  \E[A(\theta, z)] = \Aprior(\theta) + 3 \sum_{h \in [\nh]} \pi_h \Alocal(\theta;h).
\end{align}
Note that in directed models, the log-partition functions depends on the parent variables (in this case, $h$).
Having recovered the marginal distribution over $\pi$, we can compute the desired expectation.
Again, this problem is convex.

\paragraph{Perturbation analysis (sketch)}
We have so far worked with exact population values $M_2,M_3,\pi,B,\mu,\theta^*$.
With finite samples, we must work the corresponding estimated quantities $\hat M_2,\hat M_3,\hat\pi,\hat B,\hat\mu,\hat\theta$.
How the error in the moments $(\hat M_2,\hat M_3)$ affects error in $(\hat\pi,\hat B)$ is well studied
(see \citet{anandkumar12moments,anandkumar13tensor}).
The latent moments $\hat\mu$ are simply a quadratic combination of $(\hat\pi,\hat B)$,
and $\hat\theta = \nabla^{-1} A(\hat\mu)$, which is Lipschitz.
Therefore, we have a consistent estimator with polynomial sample complexity.