### 2013-05-31 doc/nips2013/discussion.tex
spell check, mods to experiments

\section{Discussion}

In recent years, there has been much interest
on using the method of moments and spectral decomposition
for learning latent-variable models.
One line of work uses observable operator models to learn
HMMs \cite{hsu09spectral},
low-treewidth graphs \cite{parikh12spectral},
probabilistic context-free grammars \cite{cohen12pcfg},
weighted finite state automata \cite{balle12automata},
finite state transducers \cite{balle11transducer},
but do not produce parameter estimates.

% Method of moments
Our work builds on a second line of work which focuses
on parameter estimation.  These methods have been used to learn
hidden Markov models \cite{anandkumar12moments},
Latent Dirichlet Allocation \cite{anandkumar12lda},
linear Bayesian networks \cite{anandkumar12linear},
mixture of spherical Gaussians \cite{hsu13spherical},
latent parse tree models \cite{hsu12identifiability}, and
mixture of linear regressions \cite{chaganty13regression}.
Not only can actual parameter estimates be valuable for scientific
applications, they also provide a natural initialization point
for further local optimization, which has been found to improve accuracy
\cite{chaganty13regression}.

Extending this latter perspective,
our work departs from the standard goal of
insisting on constructing consistent parameter estimators, but rather strives
to use spectral decomposition to obtain constraints to improve
parameter estimation.  Using these techniques,
we found empirically that our approach provides
improvements in various models such as the grid model and factorial HMM
for which we do not have a consistent parameter estimate.
We believe using partial information from factorized moments
in conjunction with likelihood optimization
broadens the toolbox of techniques for learning
latent-variable models.
Finally, our technique still relies fundamentally on
having three-view bottlenecks \cite{anandkumar12moments}.
How to extract partial information without this assumption
remains an interesting open problem.

% Structure recovery
%structure recovery
%\cite{anandkumar11tree}

% Alternate estimators
%There are multiple consistent estimators.
%Composite likelihood \cite{besag75pseudo,lindsay88composite,liang08asymptotics}