### 2013-05-31 doc/nips2013/appendix.tex
a

\appendix

\section{Directed models}
\label{sec:directedModels}

We have so far focused on undirected models for simplicity,
but our techniques also apply to directed models in which each
local conditional distribution is a log-linear model.

The setting is as follows:~Assume we have a directed acyclic graph over
nodes $V$,
and define $\pi(j) \subset V$ to be the parents of each node $j \in V$.
Define the joint distribution as a product of locally-normalized models:
\begin{align}
  p_\theta(z) = \prod_{j \in V} p_\theta(z_j \mid z_{\pi(j)}), \quad\quad
  p_\theta(z_j \mid z_{\pi(j)}) = \exp \{ \phi_j(z_j, z_{\pi(j)})^\top\theta - A_j(\theta; z_{\pi(j)}) \}.
\end{align}
Here, each local model has its own local features $\phi_j(z_j, z_{\pi(j)})$
and its own local log-partition function $A(\theta; z_{\pi(j)})$, which depends on values of its parent.
To consolidate notation with \refeqn{undirectedSetup},
define the sum of all the feature vectors
$\phi(z) = \sum_{j \in V} \phi(z_j, z_{\pi(j)})$
and $A(\theta; z) = \sum_{j \in V} A_j(\theta; z_{\pi(j)})$,
so that:
\begin{align}
  p_\theta(z) \eqdef \exp\{ \phi(z)^\top\theta - A(\theta; z) \}.
\end{align}

\subsection{Directed mixture model}

In this section, we demonstrate parameter estimation for the directed mixture
model (\reffig{directedModels}(a)).  This is the usual mixture model, defined
as follows:
\begin{align}
p_\theta(x,h) = p_\theta(h) \prod_{j=1}^3 p_\theta(x_j \mid h),
\end{align}
where
$p_\theta(h) \eqdef \exp\{\phiprior(h)^\top\theta - \Aprior(\theta)\}$
and
$p_\theta(x_j \mid h) \eqdef \exp\{\philocal(x_j, h)^\top\theta - \Alocal(\theta; h)\}$.

Consequently, the joint log-likelihood is:
\begin{align}
  L(\theta) = {\underbrace{\E[\phi(x,h)]}_{\mu}}^\top\theta - \E[A(\theta; z)], \quad
  \E[A(\theta, z)] = \Aprior(\theta) + 3 \sum_{h \in [\nh]} \pi_h \Alocal(\theta;h).
\end{align}

Note that in directed models, the log-partition functions depends on the parent variables (in this case, $h$).
Having recovered the marginal distribution over $\pi$, we can compute the desired expectation.
Again, this problem is convex.

\Fig{figures/directedModels}{0.3}{directedModels}{
Various directed versions of the undirected models.
}

\Fig{figures/factorialModels}{0.3}{factorialModels}{
Directed factorial models.
}

\subsection{Other models}

\reffig{directedModels} shows examples of the directed models
for which we have already derived undirected estimates.
In general, the aggregation
and factorization steps (Steps 1 and 2) in \refsec{threeViewMixtureModel}
remain the same as the undirected cases since these steps
only rely on the conditional independence structure of the models,
not particular parametrizations.

\subsection{Factorial models}

For factorial models, the directed and undirected version do differ in their
conditional independence statements.
For example, consider the 
the directed factorial mixture model,
also known as a sigmoid belief network \citep{saul96sigmoid},
in \reffig{factorialModels}(a).
Here ,$h_1$ and $h_2$ are marginally independent in the directed
case but conditionally indepdent in the undirected case.
However, since our algorithm treats $(h_1,h_2)$ as a composite variable $c$
anyway, it is not sensitive to these differences.
The factorial HMM \cite{ghahramani97fhmm} is similar and can be handled
similar to our treatment of HMMs plus unshuffling factorization.

\subsection{Experiments}

Table~\ref{tab:errorsa} shows the full set of results for the mixture
model and HMM.

\begin{table}
\begin{center}
    \begin{tabular}{l | l l | l l l }
        Algorithm & $k$ & $\ell$ & Accuracy & $\Delta \hat \E[\phi(x,h)]$ & Log Likelihood \\ \hline
        & \multicolumn{2}{|c|}{Mixture model} & & & \\ \hline
        \multirow{3}{*}{EM} 
        & 3 & 2 & 0.65 (+/- 0.01) & 1.23 & -3.72 \\
        & 3 & 3 & 0.51 (+/- 0.06) & 1.81 & -4.21\\
        & 5 & 2 & 0.72 (+/- 0.07) & 1.13 & -5.21\\ \hline
        \multirow{3}{*}{Spectral} 
        & 3 & 2 & 0.80 (+/- 0.09) & 0.23 & -3.54 \\
        & 3 & 3 & 0.68 (+/- 0.10) & 1.01 & -3.76 \\
        & 5 & 2 & 0.80 (+/- 0.05) & 0.18 & -5.04 \\ \hline
        \multirow{3}{*}{Oracle} 
        & 3 & 2 & 0.80 (+/- 0.08) & 0.01& -3.54 \\
        & 3 & 3 & 0.70 (+/- 0.11) & 0.02& -3.87 \\
        & 5 & 2 & 0.80 (+/- 0.05) & 0.01& -5.03 \\ \hline
        & \multicolumn{2}{|c|}{HMM} & & & \\ \hline
        \multirow{3}{*}{EM} 
& 3 & 2 & 0.64 (+/- 0.10) & 1.71 & -4.51 \\  
& 3 & 3 & 0.42 (+/- 0.05) & 2.41 & -6.09 \\  
& 5 & 2 & 0.62 (+/- 0.09) & 1.60 & -6.47 \\  \hline
        \multirow{3}{*}{Spectral} 
& 3 & 2 & 0.67 (+/- 0.07) & 2.59 & -3.71 \\ 
& 3 & 3 & 0.45 (+/- 0.08) & 4.96 & -4.36 \\ 
& 5 & 2 & 0.64 (+/- 0.10) & 3.43 & -5.31 \\ \hline
        \multirow{3}{*}{Oracle} 
& 3 & 2 & 0.71 (+/- 0.07) & 0.39 & -4.64 \\ 
& 3 & 3 & 0.47 (+/- 0.07) & 1.53 & -5.85 \\ 
& 5 & 2 & 0.73 (+/- 0.07) & 0.39 & -6.16 \\ \hline
    \end{tabular}
    \caption{Micro-averaged results for various models, methods.}
    \label{tab:errorsa}
\end{center}
\end{table}