### 2014-01-18 doc/local-optima/paper.tex
Updated proof

\documentclass[tablecaption=bottom]{jmlr}

\jmlrproceedings{}{}
\jmlrpages{}{}
\usepackage[cm]{fullpage}
\usepackage{booktabs}
\usepackage{ctable}

\title{Counting local-optima in a log-linear model}

\author{Arun Tejasvi Chaganty \Email{chaganty@cs.stanford.edu}}

\input{macros}

\begin{document}

\maketitle

\section{Introduction}

At the high-level, we're looking to count the number of the local optima
of a log-linear model,
\begin{align}
p_{\theta}(x) &= \sum_z \exp(\theta^T \phi(x,z) - A(\theta)).
\end{align}

\todo{Describe mixtures of Gaussians}

\section{Preliminaries}

\section{Mixture of Gaussians}

\section{Log-linear models}

In EM, we maximize a lower bound on the log-likelihood of model which is
derived as follows,
\begin{align}
  \sL(\theta ; \sD) 
    &\defeq \sum_{x_n \in \sD} \log p_{\theta}(x_n) \\
    &= \sum_{x_n \in \sD} \log \sum_z \exp(\theta^T \phi(x,z) - A(\theta)) \\
    &= \sum_{x_n \in \sD} \log \sum_z \frac{q(z|x_n)}{q(z|x_n)} \exp(\theta^T \phi(x,z) - A(\theta)) \\
    &= \sum_{x_n \in \sD} \sum_z q(z|x_n) \log \frac{1}{q(z|x_n)} \exp(\theta^T \phi(x,z) - A(\theta)) \\
    &\le \sum_{x_n \in \sD} \sum_z q(z|x_n) [\theta^T \phi(x,z) - A(\theta) - \log q(z|x_n)].
\end{align}

We use the notation $\sL(\theta, q) \defeq \sum_{x_n \in \sD} \sum_z
q(z|x_n) [\theta^T \phi(x,z) - A(\theta) - \log q(z|x_n)]$. The solution
to EM is $\max_\theta \max_q \sL(\theta,q)$ which we do via alternating
maximization between $\theta$ and $q$.

Furthermore, it can be shown that local optima of $\sL(\theta, q)$ are
also local optima of the original function, $\sL(\theta ; \sD)$.

Finally, we derive closed form solutions to the E and M steps,
\begin{align}
  q^*(\theta) 
    &\defeq \max_{q \in \Delta_{n-1}} \sL(\theta,q)\\
  \diff{\sL(\theta,q)}{q(z|x_n)} 
    &= \theta^T \phi(x_n,z) - A(\theta) - \log q(z|x_n) - 1 + \lambda \\
  q^*(z|x_n) &\propto \exp(\theta^T \phi(x_n,z) - A(\theta)) \\
  q^*(z|x_n) &= \frac{exp(\theta^T \phi(x_n,z) - A(\theta))}{\sum_{z'} exp(\theta^T \phi(x_n,z') - A(\theta))}.
\end{align}

\begin{align}
  \theta^*(q)
    &\defeq \max_{\theta \in \Re^{n}} \sL(\theta,q)\\
  \diff{\sL(\theta,q)}{\theta} 
  &= \sum_{x_n \sD}\sum_{z} q(z|x_n) \phi(x_n,z) - \diff{A(\theta)}{\theta} \\
  &= \sum_{x_n \sD}\sum_{z} q(z|x_n) \phi(x_n,z) - \E_{\theta}(\phi(x,z)) \\
  &= \hat\E_{q}[\phi(x_n,z)] - \E_{\theta}(\phi(x,z)) \\
  \E_{\theta^*}(\phi(x,z)) &= \hat\E_{q}[\phi(x_n,z)].
\end{align}

\begin{definition}(Trapping Region)
  A closed subset of $\Re^n$ is said to be trapping for $\theta \in
  \Re^n$ if for every $\theta' \in \del R$, $\sL(\theta',
  q(\theta')) \le \sL( \theta, q(\theta) )$.
\end{definition}

\begin{lemma}(Trapping property)
  For any $\theta$ and a trapping region $R$ containing $\theta$, expectation
  maximization initialized at $\theta$ converges to a fixed point in $R$.
\end{lemma}
\begin{proof}
  Let $\theta', q(\theta')$ be the new set of parameters after one
  iteration of EM. The proof proceeds by constructing a path $\sP$ from
  $\theta$ to $\theta'$ such that every point along the path
  $\theta^\alpha \in \sP$ has a higher likelihood than $\theta$ and thus
  could not cross the boundary $\del R$.

  Let us define a trajectory $\sP : \Re \to \Re^n$ from $\theta$ to
  $\theta'$ as follows,
  \begin{align}
    \sP(\alpha) &\defeq \alpha \theta + (1-\alpha) \theta'.
  \end{align}
  For notational convenience, we use $\theta_{\alpha} \defeq \sP(\alpha)$.

  Now, note that with $q$ fixed, $\sL(\theta,q)$ is {\em convex} in $\theta$. Thus, 
  $\sL(\theta^\alpha, q) = (1-\alpha)\sL(\theta, q) + \alpha
  \sL(\theta', q) > \sL(\theta, q)$. Furthermore, $\sL(\theta^\alpha,
  q^*(\theta^\alpha)) \ge \sL(\theta^\alpha, q) > \sL(\theta, q)$.

  \todo{Refine with a path for $q$ too}
\end{proof}

Next, we will define a local region around some parameter $\theta$ that is trapping.

\end{document}