### 2014-01-26 doc/icml2014/measurements.tex
Updated paper

\section{Parameter Estimation for Log-linear Latent-Variable Models}
\label{sec:log-linear}

\paragraph{Setup}
Let $z$ be a collection of variables indexed by $V$ (that is, $z = \{z_j\}_{j \in V}$).
Suppose that some of the variables are observed ($x = \{ z_j \}_{j \in O}$, where $O \subset V$) 
and the rest are latent ($h = \{ z_j \}_{j \in H}$, where $H = V \backslash O$).
Let $C \subset 2^V$ be a collection of subsets of variables,
and let $\phi_c(z_c) \in \Re^d$ be the feature map for each $c \in C$.
In this paper, we will consider both directed and undirected models over $z$.
In the undirected case, we define the standard globally-normalized exponential family model:
\begin{align}
  \label{eqn:undirectedSetup}
  p_\theta(z) = \exp\{ \phi(z)^\top\theta - A(\theta) \},
\end{align}
where $\phi(z) \eqdef \sum_{c \in C} \phi_c(z_c) \in \Re^d$ is the global feature vector, $\theta \in \Re^d$ is the parameter vector,
and $A(\theta) = \log \int \exp\{\phi(z)^\top\theta\} dz$ is the global log-partition function.
To some extent, our methods only depend on the conditional independence
structure of the models.

\paragraph{Problem statement}

% Statement
This paper focuses on the problem of parameter estimation:
We are given $n$ i.i.d.~examples of the observed variables $D = (x^{(1)}, \dots, x^{(n)})$
where each $x^{(i)} \sim p_{\theta^*}$ for some true parameters $\theta^*$.
Our goal is to produce a parameter estimate $\hat\theta$ that approximates $\theta^*$.

% Maximum likelihood
The standard estimation procedure is maximum (marginal) likelihood,
$\sL \eqdef \max_{\theta \in \Re^d} \sum_{x \in D} \log p_\theta(x)$,
which is statistically efficient but computationally intractable.
In practice, one uses gradient-based optimization procedures (e.g., EM or L-BFGS)
on the marginal likelihood, which can get stuck in local optima.

\paragraph{Primer: Three-view log-linear mixture models}

\citet{anandkumar12moments} describe an algorithm to recover parameters
  for a directed three-view mixture model. 
We will consider an undirected log-linear model variant of this model,
  shown in \figureref{three-view}.
The model has features $\phi(x,y) = \{\bI(x_1,y_1), \bI(x_2,y_1), \bI(x_2,y_1)\}$.

Recover moments with tensor factorization.
Show that the constrained optimization problem is convex.

Statement: consistency.

\paragraph{General latent-variable models with bottlenecks}

Our learning algorithm for three-view mixture models relies on the fact that
the sufficient statistics of the exponential family
can be expressed as a function of the latent moments.
Of course, these conditions do not hold for general log-linear models.
Rather than falling back on plain local optimization of the likelihood in these
cases, we would still like
to somehow use the \emph{partial information} offered by the latent moments.

Partial information is available when the following holds:
%\begin{property}
%  There exists some latent variable $h_{j_0} \in [\nh]$ (the bottleneck)
%  and three conditionally independent observed variables $x_{j_1},x_{j_2},x_{j_3}$ (views) such that
%  each conditional mean matrix $\E[\phix(x_{j}) \mid h_{j_0}] \in \R^{\nphix \times \nh}$ ($j \in \{ j_1, j_2, j_3 \}$) has full column rank $\nh$
%  for some transformation $\phix$.
%\end{property}

Optimization.