### 2014-01-30 doc/icml2014/related.tex
Updated undirected section

\section{Related work}

Anandkumar - linear bayesian nets

Le song - trees

David sontag - noisy or