### 2014-01-31 doc/icml2014/abstract.tex
Updated draft

\begin{abstract}

Parameter estimation in latent variable models is prone to local
  optima due to the non-convexity of the likelihood. 
In this paper, we extend recent advances in the method of moments 
  to provide consistent parameter estimates for a general class of
  directed and undirected graphical models, thus avoiding the issue of
  local optima.
In particular, our algorithm applies to any graphical model wherein each
  hidden variable has at least three conditionally independent observed
  variables.
This family includes loopy graphs with high treewidth.
We use the method of moments to tie each hidden to some observed
  variables, and show that the rest of the problem can be cast as convex
  optimization of the likelihood.
This approach gracefully extends to models outside this family by
  providing partial information via posterior regulraization.
\end{abstract}

\paragraph{Last updated:} \today, \currenttime

\paragraph{Errata / Incomplete}
\begin{enumerate}
  \item Use \texttt{\textbackslash left(}.
\end{enumerate}