### 2014-01-31 doc/icml2014/examples.tex
move examples

%\subsection{More examples}

%We will now instantiate our algorithm for a couple more examples, illustrated in \figureref{examples}.

\paragraph{Aggregating observations}
We describe two practical considerations to use samples more efficiently.
Firstly, if we have multiple exclusive views for a hidden variable,
  intuitively, it is better to aggregate over them. 
For example, consider a hidden variable $h_1$ with multiple exclusive
  views $x_1$ and $x_2$.
With the method of moments perspective, to learn the marginal
  distribution over $h_1$, one must solve the following reconstruction
  problem, 
\begin{align*}
  \hat Z_{h_1} &= \arg\min_{Z_{h_1}} \half \|Z_{h_1}(\mOpp{1}{1}) - M_1 \|^2 + \half \|Z_{h_1}(\mOpp{2}{1}) - M_2 \|^2,
\end{align*}
which does not have a closed form solution. 
In contrast, this naturally fits into the convex optimization framework, where $\hat Z_{h_1}$ will now be,
\begin{align*}
  \hat Z_{h_1} &= \arg\min_{Z_{h_1}} \sum_{\vx \in \sD} \log Z_{h_1}( \mOpp{1}{1}[x_1] \odot \mOpp{1}{1}[x_1] ),
\end{align*}
where $\cdot$ denotes element-wise multiplication.
An important note to make is that \TensorFactorize only returns
  a solution up to permutation; if $\mOpp{1}{1}$ and
  $\mOpp{1}{2}$ do not belong to the same bottleneck (e.g. $x^a_2$ and
  $x^b_2$ in \figureref{examples-tree}), then some
  care must be taken to ensure they have the same labelling.

Secondly, we can exploit parameter sharing by aggregating over
  disjoint sets of observed variables. 
For example, in \figureref{examples-hmm}, we can aggregate the statistics for
  bottlenecks $h_i$ with views $\{x_{i-1}, x_{i}, x_{i+1}\}$ before
  running \TensorFactorize; this will give us a consistent estimate for
  $O$ (as well as $T$).