### 2014-01-31 doc/icml2014/abstract.tex
push

\begin{abstract}

Recent work in method of moments provide consistent estimates
  for latent-variable models, avoiding local optima issues,
  but these methods can only be applied to certain types of graphical models.
In this work, we show that the method of moments in conjunction
  with a composite marginal likelihood objective
  yields consistent parameter estimates for a much broader
  class of directed and undirected graphical models,
  including loopy graphs with high treewidth.
%In particular, our algorithm applies to any graphical model wherein each
  %hidden variable has at least three conditionally independent observed
  %variables.
  %that sidestep local optima issues for a general class of
  %directed and undirected graphical models.
%Parameter estimation in latent-variable models based on EM
  %is prone to local optima.
  %due to the non-convexity of the marginal log-likelihood. 
%In this paper, we extend recent advances in the method of moments 
  %to provide consistent parameter estimates
  %that sidestep local optima issues for a general class of
  %directed and undirected graphical models.
%In particular, our algorithm applies to any graphical model wherein each
  %hidden variable has at least three conditionally independent observed
  %variables.
%This family includes loopy graphs with high treewidth.
Specifically, we use tensor factorization to reveal partial information
about the hidden variables, rendering the otherwise non-convex negative
log-likelihood convex.
%We use the method of moments to tie each hidden to some observed
  %variables, and show that the rest of the problem can be cast as convex
  %optimization of the likelihood.
Our approach gracefully extends to models outside our class by
  incorporating the partial information via posterior regulraization.
\end{abstract}

%\paragraph{Last updated:} \today, \currenttime
%
%\paragraph{Errata / Incomplete}
%\begin{enumerate}
%  \item Use \texttt{\textbackslash left(}.
%  \item Use $\ell$ instead of $L$.
%  \item convex and negative log likelihood.
%\end{enumerate}
%