### 2014-01-31 doc/icml2014/discussion.tex

\section{Discussion}
\label{sec:discussion}

% Summarize paper
For latent-variable models,
there exists a tension between
local optimization of likelihood,
which is broadly applicable but has no theoretical guarantees,
and the method of moments, which provides consistent estimators
but are limited to models with special structure.
The purpose of this work is to show that the two methods
can be used synergistically to produce consistent estimates
for a broader class of models.
The key idea is to (i) perform tensor factorization \citep{anandkumar13tensor} to
obtain information about each hidden variable in the form of conditional moments;
(ii) use composite likelihood to estimate the marginal distributions of latent variables;
and (iii) in the case of undirected models,
use the marginal distributions to formulate a convex optimization to
recover the model parameters.

% Bottleneck is key
Our approach allows us to handle a broad class of directed and undirected
models in which each hidden variable is a ``bottleneck''---that is,
it has three conditionally independent observations,
(although the dependence on the observations could be quite complex as a function of the parameters).
This bottleneck property of \citep{anandkumar13tensor}
has been exploited in many other contexts,
including Latent Dirichlet Allocation \cite{anandkumar12lda},
parsing \cite{hsu12identifiability},
noisy-or Bayesian networks \cite{halpern13noisyor},
mixture of linear regressions \cite{chaganty13regression},
and others.

%Much recent work on method of moments for estimating latent-variable models
%has focused on increasing applicability to broader range of models,

%In this paper, we we've described an algorithm that can consistently
%  estimate parameters for the class of directed and undirected graphical
%  models for which each hidden variable has at least three conditionally
%  independent observed variables (\propertyref{bottleneck}).

%This condition is sufficiently general to capture a variety of popular
  %models like latent trees and grids (like a pairwise latent Markov random field).
%The undirected model setting applies to log-linear models, allowing
%  for discriminative featurization; furthermore, in practice, the
%  bottleneck condition can be enforced on a model if the observed
%  features can be split into two uncorrelated sets.

% Classes of models
Notably, the bottleneck property allows us to perform consistent parameter
estimation in undirected log-linear models with hidden variables with high
treewidth, which

The key property we exploited was the presence of exclusive views
  (\propertyref{exclusive-views}) with learnable parameters.
In some sense, the presence of these views helps tie down the hidden
  variables to the data. 
While other authors, e.g \citet{anandkumar12lda, anandkumar2013linear,
  halpern13noisyor} are able to learn parameters for models outside this
  family, they do so with some knowledge of the correlation structure
  between variables. 
In a pre-processing step, they ``subtract'' out the correlations and
  recover a graph that satisfies the bottleneck property
  \propertyref{bottleneck}.
% TOO BOLD
%Thus, while \propertyref{bottleneck} is much stronger than
%  \propertyref{exclusive-views}, we expect that without any knowledge of
%  the correlation between variables, \propertyref{bottleneck} is
%  necessary for identifiability.

% Broad interest in learning more models
The use of method of moments for learning latent-variable models
%\cite{balle11transducer}
including 
Much of the Parameter estimation for latent-variable models using the method of
  moments has received much attention in recent years,
aimed at increasing their applicability to broader classes of models.

Our work uses the \TensorFactorize algorithm presented by
  \citet{anandkumar13tensor}, first
  studied the consistent parameter estimation algorithm for the
  three-view mixture model.

%\todo{What about Mossel and Roch?}

\citet{anandkumar12lda} are able to recover parameters for correlated
  discrete mixtures, e.g. latent Dirichlet allocation; their approach
  follows that of the three-view mixture model, with a pre-processing step
  that subtracts the correlation between the views.
\citet{anandkumar2013linear} study a special family of Bayesian networks
  in which the observed variables are linearly related to a sparse set of
  hidden variables. Once again, the key step is to identify the
  correlations between observed variables and exclude them from the views.
\citet{halpern13noisyor} propose an algorithm to recover factors for
  a more densely connected bipartite noisy-or network; they also use
  bottlenecks to learn parameters, but exploit properties of the model
  to negate correlations from learned edges, allowing for more
  bottlenecks to be discovered.

work to settings that do not meet the
  criterion \propertyref{bottleneck}, but rely on knowing more about
  the underlying correlation structure of the problem.

In comparison, our work applies to a general family of models, including
  latent tree models, directed and undirected grids, with no assumptions
  on the correlation structure aside from independences encoded in the
  graph.

Another prominent line of work in the method of moments community has
  focussed on recovering parameters for a transformation that admits an {\em observable
  representation}, i.e. one whose marginals can be directly computed from the data.
\citet{song2011spectral} propose an algorithm to learn an observable
  operator representations for latent tree graphical models, like the
  one in \figureref{examples-tree}, assuming \propertyref{bottleneck}. 
Their approach is similar to ours, but restricted to trees with observed
  variables at the leaves.
\citet{parikh12spectral} extend this approach to general graphical
  models in which \propertyref{bottleneck} holds, but consider the latent
  junction tree representation. 
Consequently, the observable representations have size atleast that of
  a node in the junction tree, which will be of the order of the treewidth
  of the graph. 
Our algorithm only constructs moments of the order of the cliques in the
  graph, which can be exponentially smaller. 
For example, an $n\times n$ grid model has a treewidth of $n$, but each
  clique is of size at most $2$ in the undirected case and $3$ in the
  directed case.
In some sense, an observable operator re-formulation of our algorithm
  subsumes this work.

% Identifiability
Of course, there are models which 
\cite{kruskal77three}