### 2014-03-18 doc/icml2014/rebuttal.txt
Added rebuttal

Thank you very much for your constructive comments and suggestions; we
will try to respond to your questions below:

# (All reviewers) Regarding empirical comparisons

We are actively looking into studying the empirical performance of our
proposed algorithms.

# (Reviewer 1) Regarding the canonical transformation in Lemma 1

- The parameter space defined after the canonical transformation (Lemma
  1) is only artificially larger in that it just simplifies notation
  without actually changing the underlying learning problem. 
- In particular, when the conditional moments $O^{(v,new}) = P(x_v
  | h_{new})$ are desired, we already know them to be the identity, $I$,
  and do not need to use the tensor factorization method to learn them.
- Furthermore, in the case when the conditional moments for another
  hidden variable (say $h_1$) are desired, e.g. $O^{(v,1}) = P(x_v
  | h_1)$, then only a single $x_v$ need be used. Thus, the
  transformation does not actually increase the size of the
  parameter space.
- Finally, the identity transformation between $x_v$ and $h_new$ lets us
  write $p(x_v | h_1, h_2) = p(h_new | h_1, h_2)$.

# (Reviwer 3) Novelty in the directed graphical model case

We agree that our work presents a natural extension of the work of
Anandkumar et al., however, we believe our method still offers two novel
contributions in the directed case:
- Firstly, we show that directed graphical models with high tree width
  can be _efficiently_ learned using the spectral method. Though Parikh
  et al. (2012) have studied recovering parameters in a similar class of
  graphical models in the observable operator setting, the
  method presented there relies on a junction tree representation that
  scales with the treewidth. The algorithm we present scales with
  the degree of each variable.
- Secondly, we have shown that using composite likelihoods leads to
  asymptotically better estimators; this result applies in the directed
  setting as well.

# (Reviwer 3) Extensions to more complex distributions

Given the three view assumption, our work straightforwardly generalizes
to the case where the observed variables take real values. Following
similar treatment as in Anandkumar et al. 2013a, we would be able to
recover the conditional means of the output variables. Techniques like
those proposed in Anandkumar et al. (2012a) or Hsu and Kakade (2012)
could be extended to recover covariances in limited settings (i.e.
Dirichlet distributions and spherical Gaussians respectively), but the
problem remains hard in the general case.