### 2014-05-07 doc/icml2014/assumption-proof.tex
Included section about tensor projection and how to get Y in the assumption proof

\subsection{Recovering conditional moments}
\label{app:assumption-proof}

% Define up front that we will focus on h, x_1, x_2, x_3
In step 1 of \LearnMarginals, we used the bottleneck property of a hidden
  variable $h_i$ to learn conditional moments $\mOpp{v}{i} \eqdef
  \Pr(x_v | h_i)$ for every view $x_v \in \sV_{h_i}$ using
  \TensorFactorize. 
In order to do so, we require that \assumptionref{full-rank} holds, i.e.
\begin{assumption*}
  Given a bottleneck $h_1$ with views $x_1, x_2, x_3$, the conditional
  moments $\mOpp{1}{1}, \mOpp{2}{1}, \mOpp{3}{1}$ have full column rank
  $k$, and $\pi\oft 1 \succ 0$.
\end{assumption*}

To interpret this assumption better, we would like find an equivalent
condition on the model parameters $Z_\sC$ that imply the above condition
on $\mOpp{v}{i}$. In this section, we will show that
\assumptionref{full-rank} is actually implied by the following condition
(\assumptionref{full-rank-plus});
\begin{assumption*}
For every clique $\sC \in \sG$ (including ones involving observed
  variables) and for every {\em conditioning} of the clique's marginals,
  every unfolding has full rank. 
\end{assumption*}

% Tour guide
Going forward, we will describe $\mOpp{v}{i}$ recursively in terms of
  tensor multiplications of conditional distributions $\Pr(h_c \given
  \Pa(h_c))$ in a variable elimination ordering.
We then show that the above assumption implies the unfoldings of the
  conditional distributions are full rank as well.
The analysis also tracks singular values, allowing us to present
  a sample complexity bound that highlights an exponential dependence on
  the size of the variable elimination set.

\paragraph{A recursive construction of $\mOpp{v}{i}$}

Without loss of generality, let $i = 1$. We will express the conditional
distribution $\mOpp{v}{1} \eqdef \Pr(x_v | h_1)$ recursively through the
parents of $x_v$, its parents, and so on. The approach is analogous to
message passing, and the key idea is to express the same in the form of
a tensor multiplication. \figureref{message-proof} outlines the
procedure.

\begin{figure}[t]
  \label{fig:message-proof}
  \centering
  \subimport{figures/}{assumption.tikz}
  \caption{A recursive construction of the conditional moments $\mOpp{v}{i}$. Shaded regions indicate participating variables in the intermediate distributions.}
\end{figure}

Recall that $x_v$ is the observed variable in consideration; let $h_t$
be its unique parent. Then, 
\begin{align}
  \mOpp{v}{1} &\eqdef \Pr( x_v \given h_1 ) \nonumber \\
              &= \sum_{h_t}  \Pr( h_t \given h_1 ) \Pr( x_v \given h_t ) \nonumber\\
              &= \mOpp{v}{t} \mYpp{t}{1}, \label{eqn:expanding-O}
\end{align}
where $\mYpp{i}{j} \eqdef \Pr( h_i \given h_j )$. 
More generally, for two sets of hidden variables $C \eqdef \{h_{C_1}
\cdots h_{C_m} \}$ and $C' \eqdef \{h_{C'_1} \cdots h_{C'_n} \}$, 
define $\mYpp{C}{C'} \eqdef \Pr( h_{C_1} \cdots h_{C_m} \given h_{C'_1}
\cdots h_{C'_n} )$. 
With this representation, $\mYpp{C}{C'}$ is a $|C| + |C'|$-th order
tensor.

Going forward, we will construct $\mYpp{v}{1}$ recursively.
Without loss of generality, let $h_1, h_2, \cdots, h_t$ be a topological
  ordering rooted at $h_1$,
and let $\Pa(h)$ be the parents of a hidden variable $h$ in
  this topological ordering.
Then $\mYpp{t}{1}$ can be expressed as,
\begin{align*}
  \mYpp{t}{1} &\eqdef \Pr( h_t \given h_1 )  \\
  &= \sum_{\vh \in H_{\Pa(h_t)}} \Pr( \Pa(h_t) \given h_1 ) \Pr( h_t \given \Pa(h_t) ) \\
  &= \mYpp{t}{\Pa(h_t)} \times_{\Pa(h_t)} \mYpp{ \Pa(h_t) }{1},
\end{align*}
where $A \times_{C} B$ refers to summation along the indices $C$. We
refer to the operation $A \times_C B$ as tensor multiplication.
\appendixref{tensor-multiplication} proves some key properties of this
operation.

Note that $\mYpp{t}{\Pa(h_t)}$ is the conditional probability $\Pr(h_t
 | \Pa(h_t))$, which is assumed to be full rank, and can be easily
 computed from the hidden marginals $Z_\sC$ that contain $h_t$.

Now, let's further expand $\mYpp{ \Pa(h_t) }{1}$ in terms of {\em its
  parents}, recursively doing so until we reach $h_1$.
Let $C$ be an intermediate set of hidden variables. Let $h_c \neq h_1$
  be some hidden variable in $C$ which we wish to eliminate. 
Let $C' = C \union \Pa(h_c) \setminus \{h_c\}$, the new set of hidden
  variables containing the parents of $h_c$ instead of $h_c$ and $\del
  C = C' \setminus (C \union \{h_1\})$, the variables in the interface
  between $C$ and $C'$.
Note that $\Pa(h_c) \subseteq \del C$.
Then,
\begin{align}
  \mYpp{C}{1} &\eqdef \Pr( h_c \given h_1 ) \nonumber \\
  &= \sum_{\vh \in H_{\del C}} \Pr( h_c \given \Pa(h_c) ) \Pr( C' \given h_1 ) \nonumber \\
  &= \mYpp{c}{\Pa(h_c)}  \times_{\del C} \mYpp{ C' }{ 1 }. \label{eqn:recursive-step}
\end{align}
This process is repeated until the base case, $\mYpp{1}{1} = \ones$.
\algorithmref{Y} summarizes the procedure.

\begin{algorithm}
  \caption{$\mYpp{C}{1}$}
  \label{algo:Y}
  \begin{algorithmic}
    \REQUIRE The root $h_1$, a set of hidden variables $C$.
    \ENSURE The hidden moments distribution $\mYpp{C}{1}$.
    \IF{ $C = \{h_1\}$ }
      \STATE $\mYpp{1}{1} = \ones$.
    \ELSE
      \STATE Let $h_c \neq h_1$ be some hidden variable in $C$.
      \STATE Let $C' = C \union \Pa(h_c) \setminus \{h_c\}$.
      \STATE $\mYpp{C}{1} = \mYpp{C'}{1} \times_{\Pa(h_c)} \mYpp{c}{\Pa(h_c)}$.
    \ENDIF
  \end{algorithmic}
\end{algorithm}

\paragraph{Rank conditions for $\mYpp{c}{\Pa(h_c)}$}

\algorithmref{Y} constructs an expression for $\mYpp{t}{1}$ entirely as
  tensor multiplications of terms of the form $\mYpp{c}{\Pa(h_c)}$ for
  some hidden variable $h_c$. 
The singular values of unfoldings of individual $\mYpp{c}{\Pa(h_c)}$ can
  be easily computed from the clique marginals $Z_\sC$ as follows.

\begin{lemma} 
  Let $\sC_0 \eqdef {h_c} \union \Pa(h_c) \subseteq \sC_1 \union \cdots
  \union \sC_\ell$ for $\sC_1, \cdots, \sC_\ell \in \sG$. Note that
  $\sC_0$ need not belong in the graph.

  Then, for the unfolding $\{c\}$, we get 
  \begin{align*}
    \sigma_{\min}(\mYpp{c}{\Pa(c)}\munf{I}) &\ge \max_{i\in[\ell]}\{\sigma_{\min}(Z_{\sC_i}\munf{I_i})\},
  \end{align*}
  where $i$ is the subset of $I$ present in clique $\sC_i$.
\end{lemma} 
\begin{proof}
Without loss of generality, let $\sC \eqdef \{c\} \union \Pa(h_c)$ be a clique
  in the graph $\sG$.
If instead $\sC \subseteq \sC_1 \union \sC_2$ for some $\sC_1, \sC_2 \in
\sG$, then $Z_{\sC}$ is the marginalization of some $Z_{\sC_1 \union \sC_2}$
(which may not be a parameter in our model).

\lemmaref{tensor-projection} states that when a tensor, $Z_{\sC_1 \union
\sC_2}$, is projected by a non-zero vector $v$ (in this case, $\ones$),
then the singular values of the projection $Z_\sC$ are preserved along
any unfolding $I$ of $Z_\sC$:
\begin{align*}
  \sigma_{\min}(Z_{\sC}\munf{I}) &\ge \sigma_{\min}(Z_{\sC_1 \union \sC_2}\munf{I})v_{\min} \\
  \sigma_{\min}(Z_{\sC_1 \union \sC_2}\munf{I}) &\ge
  \sigma_{\min}(Z_{\sC}\munf{I})v_{\max}\inv.
\end{align*}

In this case, we get that $\sigma_{\min}(Z_{\sC_1 \union \sC_2}\munf{I})
= \sigma_{\min}(Z_{\sC}\munf{I})v_{\max}\inv$.

Furthermore, assuming that any unfolding of $Z_{\sC_1}$ is full rank, we
have that $Z_{\sC_1 \union \sC_2}$ is also full rank along any common
unfolding.

Proceeding,
\begin{align*}
\mYpp{c}{\Pa(h_c)} 
  &\eqdef \Pr(h_c \given \Pa(h_c)) \\
  &= \frac{\Pr(h_c, \Pa(h_c))}{\Pr(\Pa(h_c))} \\
  \mYpp{c}{\Pa(h_c)}\munf{I}
  &= Z_\sC\munf{I} \diag{Z_\sC(\ones, \cdot, \ldots, \cdot)}^{-1},
\end{align*}
where $\diag(\sP)$ is a ``diagonal'' tensor constructed by taking the
tensor product with the identity matrix,
\begin{align*}
  \diag(\sP)[\vi,\vj] &\eqdef \sP[\vi] \delta[\vi,\vj].
\end{align*}

We have from the previous argument that $Z_\sC\munf{I}$ has full column
rank $k$; furthermore, $Z_\sC(\ones, \cdot, \ldots, \cdot)$ is
guaranteed to have at least $k$ non-zero elements because its complete
marginals, $\mPi{i} \succ 0 ~\forall h_i \in \sC$ have full support. 
Consequently, $\diag(Z_\sC(\ones, \cdot, \ldots, \cdot))$ has column rank $k$.
\todo{What happens if $\Pr(h_2, h_3) = 0$ at some point?}
\end{proof}

\paragraph{Sample complexity results}

With an expression for $\mOpp{v}{i}$, we are finally ready to describe
sufficient conditions for $\mOpp{v}{i}$ to be full rank, along with the
sample complexity for estimation.

\begin{lemma}[Sufficiency of \assumptionref{full-rank-plus}]
  \label{lem:full-rank-suff}
  Given the full rank conditions in \assumptionref{full-rank-plus}, then
  for any hidden variable $h_i$ and observed variable $x_v$,
  $\mOpp{v}{i}$ has full column rank.
\end{lemma}
\begin{proof}
Again, wlog, let $i=1$. 
The key step of the proof will be to use the property that the tensor
  multiplication in \equationref{recursive-step} preserves rank.

\theoremref{tensor-multiplication} shows that for any two tensors $A,
  B$ and an index set $C$, the unfoldings $I$ of $A \times_C B$ are full
  rank when the projected unfoldings $I_A$ of $A$ and $I_B$ of $B$ are
  full rank:
\begin{align*}
  \sigma_{\min}( (A \times_C B)\munf{I} )
    &\ge \sigma_{\min}(A\munf{I_A}) \sigma_{\min}(B\munf{I_B}).
\end{align*}

By the assumption, any unfolding of $\mYpp{c}{\Pa(h_c)}$ is full-rank. 
From \equationref{recursive-step} and
  \theoremref{tensor-multiplication}, this implies that $\mYpp{C}{1}$ is
  full rank for any $C$, including $C = \{h_t\}$.

Thus, $\mYpp{t}{1}$ is a full rank matrix. 
Again, by assumption,
  $\mOpp{v}{t}$ is full rank. From \equationref{expanding-O}, we finally
  get that $\mOpp{v}{1}$ is full rank.
\end{proof}

Next, we turn to the question of the sample complexity of estimating
$\mOpp{v}{i}$, which depends mainly on smallest singular value,
$\sigma_{\min}(\mOpp{v}{i})$.

\begin{lemma}[Singular values of $\mOpp{v}{1}$]
  \label{lem:mopp-singular-values}

Let $h_t$ be the unique parent of $x_v$, as above.
Let $h_1 \succ h_2 \cdots \succ h_t$ be a topological ordering of
  variables according to the graph $\sG$.
  Then, the smallest singular value of $\mOpp{v}{1}$ is at least the product of the smallest singular values of $\mYpp{c}{\Pa(h_c)}$.
\begin{align*}
  \sigma_{\min}(\sP) &\ge \sigma_{\min}(\mOpp{v}{t}) 
        \prod_{c \in [t]} \sigma_{\min}(\mYpp{c}{\Pa(h_c)}),
\end{align*}
where the parents $\Pa(h_c)$ are decided by the topological ordering. 
\end{lemma}
\begin{proof}
  The proof follows directly by application of
  \theoremref{tensor-multiplication} to \equationref{expanding-O} and
  \equationref{recursive-step}.
\end{proof}

Finally, using the tensor power method of \citep{anandkumar13tensor}, we
  get the following result on sample complexity.
\begin{theorem}[Sample complexity for $\mOpp{v}{i}$]
  \label{thm:sample-complexity-1}
  Wlog, let $v=1$ and $i=1$. Let $x_1, x_2, x_3$ be three views for
  $h_1$.  
  If 
\begin{align*}
  \|\hat M_{1,2} - M_{1,2}\|_{\op} &\le \epsilon & \|\hat M_{1,2,3} - M_{1,2,3}\|_{\op} &\le \epsilon,
\end{align*}
for some $\epsilon < \half$, then
with probability at least $1 - \delta$,
\begin{align*}
  \|\mOpphat{1}{1} - \mOpp{1}{1}\|_F 
    &\le  \\
    &
      O\left( k 
      \frac{{\pi\oft{1}}_{\max}/{\pi\oft{1}}_{\min}} 
      {(\sigma_{\min}(\mOpp{1}{1}) \sigma_{\min}(\mOpp{1}{2}))^{5/2}} \right) \epsilon.
\end{align*}

Furthermore, if $\sigma_{\min}(\mYpp{c}{\Pa(h_c)}) \le \sigma$ for every
  such
$\mYpp{c}{\Pa(h_c)}$, we get,
\begin{align*}
  \|\mOpphat{1}{1} - \mOpp{1}{1}\|_F 
    &\le 
      O\left( k 
      \frac{{\pi\oft{1}}_{\max}/{\pi\oft{1}}_{\min}} 
      {\sigma^{5t}} \right) \epsilon,
\end{align*}
where $t$ is the length of the topological ordering, $h_1 > \cdots
> h_t$ as defined above.
\end{theorem}
\begin{proof}
  The first statement follows directly from Theorem 5.1 of
  \citet{anandkumar13tensor} by noting that $\|M_{1,2}\|_\op \le 1$ and
  $\|M_{1,2,3}\|_\op \le 1$ (as they represent probability
  distributions). The second statement follows directly from
  \lemmaref{mopp-singular-values}.
\end{proof}