### 2014-05-12 doc/icml2014/intro.tex

\section{Introduction}
\label{sec:introduction}

% 1. Latent variable models are good.
Latent-variable graphical models provide compact representations of data
and have been employed across many fields % \todo{find a much better/broader set of citations}
\cite{ghahramani1999variational,jaakkola1999variational,blei03lda,quattoni04crf,beaumont2004bayesian,haghighi06prototype}.
% 2. Learning them is hard.
However, learning these models remains a difficult problem due to the
non-convexity of the negative log-likelihood.  Local methods such as expectation maximization (EM)
are the norm, but are susceptible to local optima.

% 3. People have approached unsupervised learning with the MoM magic sauce, but the sauce is limited.
Recently, unsupervised learning techniques based on the spectral method of moments
have offered a refreshing perspective on
this learning problem
\citep{mossel2005learning,
hsu09spectral,
bailly2010spectral,
song2011spectral,
anandkumar11tree,
anandkumar12lda,
anandkumar12moments,
hsu12identifiability,
balle12automata}.
%\todo{make sure we are citing the right people}
These methods exploit the linear algebraic properties of the model to
factorize moments of the observed data distribution into parameters,
providing strong theoretical guarantees.
However, they apply to a limited set of models, and are thus
not as broadly applicable as EM.

\begin{figure}[t]
  \label{fig:approach}
  \centering
  \subimport{figures/}{approach.tikz}
  \caption{
 % \todo{need three steps: tensor factorization, optimize composite likelihood, recover model parameters;}
  Overview of our approach:
  we first use tensor factorization to learn the \emph{conditional moments}
  for each hidden variable.
  Second, we optimize a composite marginal likelihood to recover the \emph{hidden clique marginals},
  from which we solve for the \emph{model parameters}.
  }
\end{figure}

% DONE: need to say from the beginning that we have a hybrid approach, not just using method of moments;
% there are really two contributions: one is the consistent estimate result for class of models,
% and the other is the general idea of using partial constraints]

% 4. State what we do: exploit moment constraints to make the problem easier.
In this paper,
we show that a much broader class of discrete directed and undirected graphical models can be consistently estimated:
specifically those in which \emph{each} hidden variable has three conditionally
independent observed variables (``views'')---see \propertyref{bottleneck}.
Our key idea is to leverage the method of moments,
not to directly provide a consistent parameter estimate as in previous work,
but as constraints on a likelihood-based objective.
%For a broad class of directed and undirected graphical models

%In this work, we exploit the spectral method to learn moment constraints
  %on the observed variables and show how these constraints lead to consistent parameter estimates 
  %for any graphical model with the property that there exist at least
  %three ``views'' or conditionally independent observed variables for
  %every hidden variable (\propertyref{bottleneck}).

% 5. We get moments from third-order tensors from bottlenecks and factorize them into marginals,
% then get parameters.
% DONE: try to talk about the approach in terms of three steps
The essence of our approach is illustrated in \figureref{approach}. 
First, we identify three views for each hidden variable $h_i$ (for example,
$x_1^a$, $x_1^b$ and $x_3^a$ are conditionally independent given $h_1$) and use
the tensor factorization algorithm of
\citet{anandkumar13tensor} to estimate the \emph{conditional
moments} $\Pr(x_i^a \mid h_i)$ and $\Pr(x_i^b \mid h_i)$ for each $i$ (\sectionref{directed}).
Second, we optimize a \emph{composite marginal likelihood} to recover the marginals over
each hidden clique (e.g., $\mathbb P(h_2, h_3, h_4)$).
Normally, such a marginal likelihood objective would be non-convex,
but given the conditional moments, we obtain a convex objective,
which can be globally optimized using EM (\sectionref{piecewise}).
So far, our method has relied only on the conditional independence
structure of the model and applies generically to both directed
and undirected models.
The final step of turning hidden marginals into model parameters
requires some specialization.
In the directed case, this is simple post-processing;
in the undirected case, we need to solve another convex optimization problem (\sectionref{undirected}).
%Given these factors, we show that there exists a set of observed
  %variables for every clique such that the likelihood is convex in the
  %parameters of that clique. 
%This guarantees that we can recover the true marginal distributions
  %using EM.
  %finally, with some post-processing, we are able to
  %consistently recover the true parameters of the model.
%In \sectionref{directed}, we describe our algorithm in more detail and
%show that our algorithm yields consistent parameter estimates
%under non-degenerate rank assumptions.
%describe the
  %settings and technical assumptions in which this approach is guaranteed
  %to work.
% PL: hard to talk about $O$ and $T$ at this stage.
%As a particular example, we are able to recover parameters for the
  %model depicted in \figureref{approach} if $O$ and $T$ are full-rank.

%Our approach also applies to undirected log-linear latent variable
  %models, where we use the recovered marginals to make the learning
  %problem convex (\sectionref{undirected}).
%Finally, our approach gracefully extends to more general model families
%  in which we are unable to recover the marginal distributions for all
%  hidden cliques; in such scenarios, we use posterior regularization to constrain
%  EM.  
%\todo{Is this possible?:  We show that empirically, using such constraints alleviates local
%optima issues.}