### 2014-05-13 doc/icml2014/exclusive-views.tex
Included new assumptions for conditional independence

\subsection{\lemmaref{bottleneck-views}}
\label{app:exclusive-views}

In \sectionref{directed}, we stated that bottlenecks guaranteed
exclusive views provided conditional independence assumptions \assumptionref{ci}. 
We provide the proof below.
  We generalize the lemma so that it applies
  on any connected subgraph, not just cliques.
  Also, the statement is made in terms of subgraphs,
  not the entire graph.

\begin{lemma*}[Bottlenecks guarantee exclusive views]
  %\label{lem:bottleneck-views}  
  Let $\sG$ be a graphical model satisfying conditional independence
  assumptions (\assumptionref{ci}) and be uniformly bottlenecked, then
  $\sG$ has the exclusive views property.
%Let $\sG' = \{ h_{i_1}, \ldots, h_{i_m} \} \subseteq \sG$ be
%  a connected subgraph of hidden variables. If every hidden variable
%  $h_i \in \sG'$ is a bottleneck, then $\sG'$ has the exclusive view property.
\end{lemma*}

\begin{proof}
Let $h_0$ be a hidden variable in the clique $\sC$ and $x_1, x_2$ be any
  two conditionally independent observed variables. 
One of $x_1$ or $x_2$ is an exclusive view for $h_0$ iff it is
  independent of all other hidden variables in the clique, given $h_0$.
Intuitively, if this weren't the case, then we would be able to find
  an active path between $x_1$ and $x_2$ that does not go through $h_0$,
  violating the conditional independence condition.

% TODO: One line here about how assumptions save us.
% The conditional independence assumptions (\assumptionref{ci}) exclude
%   edge cases wherein the active trail still preserves 
% for them to be both depend on any other
% variable $h'$ in the clique.

Suppose neither $x_1$ nor $x_2$ were exclusive views for $h_0$, 
  then there exists some $h_1, h_2 \in \sC$ (not necessarily distinct)  that $x_1, x_2$ are
  respectively dependent on:
  $x_1 \not\perp h_1 \given h_0$ and $x_2 \not\perp h_2 \given h_0$. 
Both $h_1$ and $h_2$ are part of the same clique $\sC$, thus by
  \assumptionref{ci}, we have that $h_1 \not\perp h_2 \given h_0$.

Since $x_1 \not\perp h_1 \given h$, there exists an active trail from
$x_1 \to h_1$. Furthermore, because $h_1$ and $h_2$ are in the same
clique, the active trail extends to $h_2$, i.e. $x_1 \not\perp h_2$.
However, since $x_1 \to h_1 \to h_2$ is an active trail in the unconditioned graph, i.e. $x_1 \not\perp h_1$; by \assumptionref{ci}, it remains active trail when conditioned on $h$.

Consequently, $x_1 \not\perp h_2 \given h_0$, which
contradicts the assumption that as $x_1 \perp x_2 \given h_0$,
$x_1 \not\perp h_i \given h_0$ and $x_2 \not\perp h_i \given h_0$ for all $h_i \in \sC$.

\end{proof}