### 2014-05-13 doc/icml2014/exclusive-views.tex
Updated exclusive views property to be true with stronger assumptions

\subsection{\lemmaref{bottleneck-views}}
\label{app:exclusive-views}

In \sectionref{directed}, we stated that bottlenecks guaranteed
exclusive views provided the graph faithfully represents conditional independences (\assumptionref{ci}). 
We provide the proof below.
  We generalize the lemma so that it applies
  on any connected subgraph, not just cliques.
  Also, the statement is made in terms of subgraphs,
  not the entire graph.

\begin{lemma*}[Bottlenecks guarantee exclusive views]
  %\label{lem:bottleneck-views}  
  Let $\sG$ be a graphical model satisfying the faithful representation assumption (\assumptionref{ci}).
  Let $\sC \in \sG$ be a uniformly bottlenecked clique, then
  $\sC$ has the exclusive views property, i.e.\ every variable in $\sC$
  has an exclusive view.
%Let $\sG' = \{ h_{i_1}, \ldots, h_{i_m} \} \subseteq \sG$ be
%  a connected subgraph of hidden variables. If every hidden variable
%  $h_i \in \sG'$ is a bottleneck, then $\sG'$ has the exclusive view property.
\end{lemma*}

\begin{proof}
By the canonical reduction (\lemmaref{reduction}), we can assume without
  loss of generality that all observed variables are leaves and thus
  all v-structures are active.
Combined with the faithful representation assumption, this lets us
  reason about independences solely in terms of graph connectivity.

Let $h_0$ be a hidden variable in the clique $\sC$ and let $x_1, x_2$ be any
  two conditionally independent observed variables. 
Recall that if $x_1$ is an exclusive view of $h_0$ in $\sC$,
  then, conditioned on $h_0$, $x_1$ is independent of every other variable $h_i
  \in \sC$.
% Since $x_1 \perp x_2 \given h_0$, every path from $x_1$ to $x_2$ must go through $h_0$. 
Suppose neither $x_1$ nor $x_2$ are exclusive views for $h_0$, 
  then there exists some $h_1, h_2 \in \sC$ that $x_1, x_2$ are respectively dependent on.
However, $h_1$ and $h_2$ are connected by participation in the clique; consequently there is a path $x_1 \to h_1 \to h_2 \to x_2$. 
By \assumptionref{ci}, this implies that $x_1$ and $x_2$ are not
  independent conditioned on $h_0$, which is a contradiction.

Thus, at least one of $x_1$ or $x_2$ is isolated from the rest of the
variables in the clique and forms an exclusive view for $h_0$ in $\sC$.

\end{proof}