### 2014-05-17 doc/icml2014/intro.tex

\section{Introduction}
\label{sec:introduction}

% 1. Latent variable models are good.
Latent-variable graphical models provide compact representations of data
and have been employed across many fields % \todo{find a much better/broader set of citations}
\cite{ghahramani1999variational,jaakkola1999variational,blei03lda,quattoni04crf,beaumont2004bayesian,haghighi06prototype}.
% 2. Learning them is hard.
However, learning these models remains a difficult problem due to the
non-convexity of the negative log-likelihood.  Local methods such as expectation maximization (EM)
are the norm, but are susceptible to local optima.

% 3. People have approached unsupervised learning with the MoM magic sauce, but the sauce is limited.
Recently, unsupervised learning techniques based on the spectral method of moments
have offered a refreshing perspective on
this learning problem
\citep{mossel2005learning,
hsu09spectral,
bailly2010spectral,
song2011spectral,
anandkumar11tree,
anandkumar12lda,
anandkumar12moments,
hsu12identifiability,
balle12automata}.
These methods exploit the linear algebraic properties of the model to
factorize moments of the observed data distribution into parameters,
providing strong theoretical guarantees.
However, they apply to a limited set of models, and are thus
not as broadly applicable as EM.

\begin{figure}[t]
  \label{fig:approach}
  \centering
  \subimport{figures/}{approach.tikz}
  \caption{
  Overview of our approach:
  we first use tensor factorization to learn the \emph{conditional moments}
  for each hidden variable.
  Second, we optimize a composite marginal likelihood to recover the \emph{hidden clique marginals},
  from which we solve for the \emph{model parameters}.
  }
\end{figure}

% DONE: need to say from the beginning that we have a hybrid approach, not just using method of moments;
% there are really two contributions: one is the consistent estimate result for class of models,
% and the other is the general idea of using partial constraints]

% 4. State what we do: exploit moment constraints to make the problem easier.
In this paper,
we show that a much broader class of discrete directed and undirected graphical
models can be consistently estimated:
specifically those in which \emph{each} hidden variable has three conditionally
independent observed variables (``views'').
Our key idea is to leverage the method of moments,
not to directly provide a consistent parameter estimate as in previous work,
but as constraints on a likelihood-based objective.
Notably, our method applies to latent undirected log-linear models with high treewidth.
%For a broad class of directed and undirected graphical models

% 5. We get moments from third-order tensors from bottlenecks and factorize them into marginals,
% then get parameters.
% DONE: try to talk about the approach in terms of three steps
The essence of our approach is illustrated in \figureref{approach},
which contains three steps.
First, we identify three views for each hidden variable $h_i$ (for example,
$x_1^a$, $x_1^b$ and $x_3^a$ are conditionally independent given $h_1$) and use
the tensor factorization algorithm of
\citet{anandkumar13tensor} to estimate the \emph{conditional
moments} $\Pr(x_i^a \mid h_i)$ and $\Pr(x_i^b \mid h_i)$ for each $i$ (\sectionref{bottlenecks}).
Second, we optimize a \emph{composite marginal likelihood} to recover the marginals over
subsets of hidden nodes (e.g., $\mathbb P(h_2, h_3, h_4)$).
Normally, such a marginal likelihood objective would be non-convex,
but given the conditional moments, we obtain a convex objective,
which can be globally optimized using EM
(see Sections~\ref{sec:exclusiveViews} and \ref{sec:piecewise}).
So far, our method has relied only on the conditional independence
structure of the model and applies generically to both directed
and undirected models.
The final step of turning hidden marginals into model parameters
requires some specialization.
In the directed case, this is simple normalization;
in the undirected case, we need to solve another convex optimization problem
(\sectionref{undirected}).