### 2014-05-17 doc/icml2014/directed.tex
simplify directed grid example

\section{Recovering hidden marginals}
\label{sec:hiddenMarginals}

\begin{figure}
  \centering
  \subfigure[Bottleneck] {
    \includegraphics[width=0.35\columnwidth]{figures/three-view.pdf}
    \label{fig:three-view}
  }
  \subfigure[Exclusive views]{
    \includegraphics[width=0.45\columnwidth]{figures/exclusive-views.pdf}
    \label{fig:exclusive-views}
  }

  \caption{(a) A bottleneck can be estimated consistently
  using tensor factorization \citep{anandkumar13tensor}.
  (b) An example of exclusive views for a bi-dependent subset $S$.
  }
\end{figure}

%In this section, we will develop a consistent parameter estimate for a class of directed graphical models.
%The conditional moments $\mOpp{v}{i} \eqdef \Pr(x_v \mid h_i)$
%recovered from {\TensorFactorize}, are not the underlying parameters.
  %Once we recover the conditional moments $\mOpp{v}{j} \eqdef \Pr(x_v
  %| h_j)$ for some $v, j$, we present a systematic approach to learn the
  %conditional probability tables for every clique for a class of directed
  %graphical models.
Having recovered conditional moments $\mOpp{v}{i} \eqdef \Pr(x_v \mid h_i)$,
we now seek to compute the marginal distribution of sets of hidden variables
$Z_S \eqdef \Pr(\bh_S)$.
%taking us one step closer to the actual model parameters.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Example: directed grid model}
%\label{sec:directedExample}

\paragraph{Example}

%The model has eight observed variables $x^a_1, x^b_1 \cdots, x^a_4, x^b_4$ and four
  %hidden variables $h_1, \ldots, h_4$.
%The parameters of this model are the conditional probability tables
%$\pi \eqdef \Pr(h_1) \in \Re^k, T \eqdef \Pr(h_2 | h_1) = \Pr(h_3 | h_1) \in \Re^{k \times k},
%V \eqdef \Pr(h_4 | h_2, h_3) \in \Re^{k \times k \times k}$ and $O \eqdef \Pr(x^a_i | h_i)
%=  \Pr(x^b_i | h_i) \in \Re^{d \times k}$.

%\paragraph{Estimating $O$}
%Since the observed variables $x^a_1, x^b_1, x^a_2$ are
%  conditionally independent given $h_1$, we can $\TensorFactorize$ from
%  \sectionref{setup} to recover $O$.

%\paragraph{Estimating $\pi$}
%The moments of $x^a_1$, $\mO_1 \eqdef \Pr(x^a_1)$ are directly related to
%  $\pi$ by a linear system: $\mO_1 = O \pi$. 
%If $O$ has full column rank, we can recover $\pi$ by taking the pseudoinverse: $\pi = O\pinv  \mO_1$.

%\paragraph{Estimating $T$}

To gain some intuition, consider the directed grid model from \figureref{approach}.
We can express the observed marginals $\mO_{12} \eqdef \Pr(x^a_1, x^a_2) \in \Re^{d \times d}$
as a linear function of the hidden marginals $\mH_{12} \eqdef \Pr(h_1, h_2) \in \Re^{k \times k}$,
where the linear coefficients are based on the conditional moments
$\mOpp{1}{1},\mOpp{2}{2} \in \Re^{d \times k}$:
%down the moments of $x^a_1, x^a_2$,
\begin{align*}
\mO_{12} = \mOpp{1}{1} \mH_{12} \mOppt{2}{2}.
\end{align*}
We can then solve for $\mH_{12}$ by matrix inversion:
\begin{align*}
\mH_{12} = \mOppi{1}{1} \mO_{12} \mOppit{2}{2}.
\end{align*}
%$T$ can be recovered from the $\mH_{12}$ by renormalizing the columns.

%\paragraph{Estimating $V$}
%Finally, we can estimate $V$ by renormalizing the hidden marginals
%$\mH_{234} \eqdef \Pr(h_2, h_3, h_4)$ from the third-order moments
%$\mO_{234} \eqdef \Pr(x^a_2, x^a_3, x^a_4)$:
%\begin{align*}
%  \mO_{234} = \mH_{234}(O, O, O) \quad\Rightarrow\quad
%  \mH_{234} = \mO_{234}(O\pinv , O\pinv , O\pinv ).
%\end{align*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Exclusive views}
\label{sec:general}

%The intuitions of the directed grid generalize readily to general graphs.
%The key property required for our algorithm to work is as follows:
For which subsets of hidden nodes can we recover the marginals?  The following
definition offers an characterization:
\begin{definition}[Exclusive views]
  \label{def:exclusive-views}
Let $S \subseteq \bh$ be a subset of hidden variables.
We say $h_i \in S$ has an exclusive view $x_v$
  if the two conditions hold:
  (i) there exists some observed variable
  $x_{v}$ which is conditionally independent of the others $S \backslash \{ h_i \}$ given $h_i$,
  and (ii) the conditional moment matrix $\mOpp{v}{i} \eqdef
  \Pr(x_{v} \mid h_i)$ have full column rank $k$ and can be recovered.
We say that $S$ has the \emph{exclusive views property} if every $h_i \in S$ has an exclusive view.
%If every hidden clique has the exclusive views property, then we say that $\sG$ does as well.
\end{definition}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Estimating hidden marginals}

We now show that if a subset of hidden variables $S$ has the exclusive views property,
then we can recover the marginal distribution $\Pr(\bh_S)$.
Consider any $S = \{i_1, \ldots, i_m\}$ with the exclusive views property. Let
  $x_{v_j}$ be an exclusive view for $h_{i_j}$ in $S$ and define $\sV
  = \{v_1, \ldots, v_m\}$. % be a set of exclusive views for the clique $S$.
By the exclusive views property,
the marginal over the observed variables $\Pr(\bx_\sV)$
factorizes according to the marginal over the hidden variables $\Pr(\bh_S)$
times the conditional moments:
\begin{align*}
  \mO_\sV 
  &\eqdef \Pr(\Sx{\sV}) \\
  &= \sum_{\vh_S} \Pr(\Sh{S}) 
                    \Pr(x_{v_1} | h_{i_1}) \cdots \Pr(x_{v_m} | h_{i_m}) \\
  &= Z_{S}(\mOpp{v_1}{i_1},\dots,\mOpp{v_m}{i_m}) \\
  &= Z_{S}(\mOppAll),
\end{align*}
where $\mOppAll = \mOpp{v_1}{i_1} \otimes \cdots \otimes \mOpp{v_m}{i_m}$ is the tensor product of
all the conditional moments.
Vectorizing, we have that
$Z_S \in \Re^{k^m}$,
$M_\sV \in \Re^{d^m}$,
and $\mOppAll \in \Re^{d^m \times k^m}$.
Since each $\mOpp{v}{i}$ has full column rank $k$,
the tensor product $\mOppAll$ has full column rank $k^m$.
Succinctly, $\mO_\sV$ (which can be estimated directly from data)
is a linear function of $Z_S$ (what we seek to recover).
We can solve for the hidden marginals $Z_S$ simply by projecting $\mO_\sV$ onto the column
space of $\mOppAll$:
\begin{align*}
  Z_{S} &= \mO_\sV(\mOpp{v_1}{i_1}\pinv,\cdots,\mOpp{v_m}{i_m}\pinv).
\end{align*}

\algorithmref{learnMarginals} summarizes the procedure, \LearnMarginals.
Given $Z_S$, the conditional probability tables for $S$ can easily be
obtained via renormalization.
\begin{theorem}[Hidden marginals from exclusive views]
If $S \subseteq \bx$ is a subset of hidden variables with the exclusive views property,
then \algorithmref{learnMarginals} recovers the marginals $Z_S = \Pr(\bh_S)$.
\end{theorem}

\begin{algorithm}
  \caption{\LearnMarginals~(pseudoinverse version)}
  \label{algo:learnMarginals}
  \begin{algorithmic}
    \REQUIRE Hidden subset $S = \{ h_{i_1}, \dots, h_{i_m} \}$ with exclusive views $\sV = \{ x_{v_1}, \dots, x_{v_m} \}$
    and conditional moments $\mOpp{v_j}{i_j} = \Pr(x_{v_i} \mid h_{i_j})$
    \ENSURE Marginals $Z_S = \Pr(\bh_S)$.
      %\STATE Identify exclusive views $\Sx{\sV} = \{x_{v_1}, \dots, x_{v_m}\}$.
      \STATE Return $\mH_S \gets \mO_{\sV}( \mOpp{v_1}{i_i}\pinv, \dots, {\mOpp{v_m}{i_m}}\pinv )$.
  \end{algorithmic}
\end{algorithm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Structural properties}

We have shown how to estimate the parameters of any clique that possesses the exclusive
  views property (\definitionref{exclusive-views}).
  But in a general graph, which cliques have this property?
  To provide intuition, we will provide more interpretable sufficient conditions
  and examples.

% Define bidependent
First, define the notion of a bidependent set (in analogy with biconnected component),
in which conditionining on one variable does not break the set apart:
\begin{definition}[Bidependent set]
Given a graphical model $\sG$, we say that a subset of nodes $S$ is \emph{bidependent} if
conditioned on any $a \in S$, there is an active trail between any other two nodes $b,c \in S$.
\end{definition}
Note that all cliques are bidependent, but bidependent sets can have more conditional independences.
This will be important for pseudolikliehood in \sectionref{pseudolikelihood}.

% bidependent, bottleneck => exclusive views
Bidependent sets are significant because they guarantee exclusive views if they are bottlenecks:
%If a subset of hidden nodes $S$ is a bottleneck, it is clear that (ii)
%of the exclusive views property (\definitionref{exclusive-views}) via \TensorFactorize,
%but we will show that part (i) follows as well.
%Now we establish the link between bottlenecks and exclusive views:
\begin{lemma}[Bottlenecked implies exclusive views]
  \label{lem:bottleneck-views}  
  Let $S \subseteq \bh$ be a bidependent subset of hidden variables.
  If $S$ is a bottleneck, then $S$ has the exclusive views property.
\end{lemma}
\begin{proof}
Let $S$ be a bidependent subset and fix any $h_0 \in S$.
Since $h_0$ is a bottleneck, it has three conditionally independent views,
say $x_1, x_2, x_3$ without loss of generality. 
For condition (i), we will show that at least one of the views is conditionally independent
of $S \backslash \{ h_0 \}$ given $h_0$.
For the sake of contradiction, 
suppose that each observed variable $x_i$ is conditionally dependent on some
$h_i \in S \backslash \{h_0\}$ given $h_0$, for $i \in \{1, 2, 3\}$.
Then conditioned on $h_0$,
there is an active trail between $h_1$ and $h_2$ because $S$ is biconnected.
Thus, there is an active trail $x_1 - h_1 - h_2 - x_2$ conditioned on $h_0$.
Since the graphical model is faithful by assumption, we have $x_1 \not\perp x_2 \mid h_0$,
contradicting the fact that $x_1$ and $x_2$ are conditionally independent given $h_0$.
To show condition (ii), assume, without loss of generality, that $x_1$ is an exclusive view.
Then we can recover $\mOpp{1}{0} = \Pr(x_1 \mid h_0)$ via $\TensorFactorize$.
\end{proof}

%Note that bottlenecks provide three views, which is important for condition (ii) of
%\definitionref{exclusive-views}, only two is required for condition (i).
% With this lemma in place, we present our full algorithm, $\LearnMarginals$,
% in \algorithmref{learnMarginals}.

% DONE: move here because not central to story
\paragraph{Remarks.}
Note that even having two independent views for each $h_i \in S$ is sufficient for condition (i)
  of the exclusive views property, while three is needed for condition (ii).
The bottleneck property (\definitionref{bottleneck} can also be relaxed if some cliques
  share parameters.
For example, in the directed grid model, we can recover the conditional moments $O$ from
  $x^a_1, x^b_1$ and $x^a_2$ with $h_1$ as the bottleneck.
  Therefore, $h_2, h_3$ and $h_4$
  need not be bottlenecks, and we can omit the observations $x^b_2, x^b_3$ and $x^b_4$.

Our method also extends directly to case in which the observed variables
  are real-valued, while the hidden variables remain discrete. 
In this setting, the tensor factorization method of
  \citet{anandkumar13tensor} recovers the expected conditional means,
  $\mu_{vi} \eqdef \E(x_v | h_i) \in \Re^{d \times k}$ for each observed variable $x_v$ and
  hidden variable $h_i$.
\algorithmref{learnMarginals} % and \algorithmref{directed} 
still applies and allow us to recover clique marginals $Z_\sC$.
%  \todo{make this a footnote if running out of space}
% PL: don't understand this comment
%In this case, we only require that the hidden variables in distinct
  %cliques share parameters.

\begin{figure}
  \centering
  \subfigure[Hidden Markov model] {
    \includegraphics[width=0.45\columnwidth]{figures/hmm.pdf}
    \label{fig:examples-hmm}
  }
%  \subfigure[Directed grid model] {
%    \label{fig:examples-grid}
%    \includegraphics{figures/grid.pdf}
%  }
  \subfigure[Tree model] {
    \includegraphics[width=0.45\columnwidth]{figures/tree.pdf}
    \label{fig:examples-tree}
  }
  \subfigure[Noisy-or model] {
    \includegraphics[height=5em]{figures/non-example.pdf}
    \label{fig:examples-noisy-or}
  }
  \caption{(a) and (b): graphs that satisfy the exclusive views property; (c) does not.}
  \label{fig:examples}
\end{figure}

\paragraph{Example: hidden Markov model.}

In the HMM (\figureref{examples-hmm}), the parameters
are $O \eqdef \Pr(x_i|h_i)$ and $T \eqdef \Pr(h_{i+1} | h_i)$
for all $i$.
While the first and last hidden variables $h_1, h_M$ in the
  sequence are not bottlenecks, they still have exclusive views ($x_1$ and
  $x_M$, respectively)
  due to parameter sharing.

\paragraph{Example: latent tree model.}

In the latent tree model (\figureref{examples-tree}), the parameters
are $\pi \eqdef \Pr(h_i)$, $T \eqdef \Pr(h_i | h_1)$ for $i \in \{2,3,4\}$,
and $O \eqdef \Pr(x^a_i | h_i) = \Pr(x^b_i | h_i)$ for $i \in \{2,3,4\}$.
Note that while $h_1$ is not directly connected to an observed variable,
  it is still a bottleneck, with views $x^a_2, x^a_3, x^a_4$.
We can recover $T$ from the clique $\{h_1, h_2\}$ by using views $x^a_2$
  (exclusive to $h_2$) and $x^a_3$ (exclusive to $h_1$).

\paragraph{Non-examples}
\label{sec:non-example}

There are certainly models which are identifiable but do not have exclusive views.
For example, \figureref{examples-noisy-or} shows
  a binary-valued noisy-or network which can be
  learned by the algorithm of \citet{halpern13noisyor},
  but does not satisfy the exclusive views property.