### 2014-05-17 doc/icml2014/undirected.tex
undirected draft finished

\section{Recovering parameters}
\label{sec:undirected}

% Transition
We have thus far shown how to recover the conditional moments
$\mOpp{v}{i} = \BP(x_v \mid h_i)$ for each exclusive view $x_v$ of each hidden variable $h_i$,
as well as the hidden marginals $Z_S = \BP(\vh_S)$ for each bidependent subset of hidden variables $S$.
Now all that remains to be done is to recover the parameters.

Since our graphical model is in canonical form (\assumptionref{canonical}),
all cliques $\sC \in \sG$
either consist of hidden variables $\vh_\sC$ or are of the form $\{x_v,h_i\}$.
The key observation is that the clique marginals are actually sufficient
statistics of the model $p_\theta$.
How we turn these clique marginals $\{ \Pr(\bx_\sC,\bh_\sC) \}_{\sC \in \sG}$
into parameters $\theta$ depends on the exact model parametrization.

For directed models, the parameters are simply the local conditional tables
$p_\theta(a \mid \Pa(a))$ for each clique $\sC = \{a\} \cup \Pa(a)$.
These conditional distributions can be obtained by simply normalizing $Z_\sC$
for each assignment of $\Pa(a)$.

%by relating the recovered conditional moments with the expected features of
%the log-linear model.

% Define model
%As before, let $\sG$ be a graphical model with observed variables $\bx$ and hidden variables $\bh$
%(see \figureref{examples-mrf} for an example).
%For each clique $\sC \in \sG$, we have a feature vector $\phi_\sC(\vx_\sC, \vh_\sC)$,
%where $\vx_\sC$ and $\vh_\sC$ are the observed and hidden variables in $\sC$, respectively.
%Let $\phi(\vx, \vh) = \sum_{\sC \in \sG} \theta^\top \phi_\sC(\vx_\sC,\vh_\sC)$ be the global feature vector.
%We consider log-linear models of the following form:
%\begin{align*}
%p_\theta(\vx, \vh) &= \exp\left( \theta^\top \phi(\vx,\vh) - A(\theta) \right),
%\end{align*}
%where $A(\theta) = \log \sum_{\vx, \vh}  \exp( \theta^\top \phi(\vx,\vh) )$ is the log-partition function.

% FIGURE: example
\begin{figure}
  \centering
  \includegraphics[width=0.6\columnwidth]{figures/mrf.pdf}
%  \subimport{figures/}{mrf.tikz}
  \caption{Example: undirected grid model where each hidden variable has two
  conditionally independent observations.
  This model has high treewidth,
  but we can estimate it efficiently using pseudolikelihood.}
  \label{fig:examples-mrf}
\end{figure}

%Of course, the log-likelihood is non-concave:
%\begin{align*}
  %L_\text{unsup}(\theta) \eqdef \E_{\vx \sim p_{\theta^*}}[\log \sum_{\vh \in \sH} p_\theta(\vx,\vh)],
%\end{align*}
%where we are assuming for simplicity that we have infinite data drawn from the true distribution $p_{\theta^*}$.
For undirected log-linear models, the canonical parameters $\theta$ cannot be obtained locally,
but we can construct a global convex optimization problem to solve for $\theta$.
Suppose we were able to observe $\vh$.
Then we could optimize the \emph{supervised} likelihood, which is concave:
\begin{align}
  L_\text{sup}(\theta)
  & \eqdef \E_{(\vx,\vh) \sim p_{\theta^*}}[\log p_\theta(\vx,\vh)] \nonumber \\
  & = \theta^\top \left(\sum_{\sC \in \sG} \E[\phi(\vx_\sC,\vh_\sC)]\right) - A(\theta).
\label{eqn:logLinearSupervised}
\end{align}
Of course we don't have supervised data,
but we do have the marginals $\Pr(\vx_\sC,\vh_\sC)$,
from which we can easily compute the expected features:
%recall that the method of moments provides the hidden marginals $Z_\sC = \BP(\vh_\sC)$
%and conditional moments $\mOpp{v}{i} = \Pr(x_v \mid h_i)$.
%Assuming our graphical model is in canonical form (\lemmaref{reduction}),
%every observed variable is connected to exactly one hidden variable.
%Then $Z_\sC$ and $\mOpp{v}{i}$ directly yield marginals for every clique $\BP(\vx_\sC, \vh_\sC)$.
%Given these marginals, we can compute the expected local feature vector:
\begin{align}
\label{eqn:logLinearFeatures}
\mu_\sC \eqdef \E[\phi(\vx_\sC,\vh_\sC)] = \sum_{\vx_\sC,\vh_\sC} \BP(\vx_\sC,\vh_\sC) \phi(\vx_\sC,\vh_\sC).
\end{align}
%Note that $\{\mu_\sC\}_{\sC \in \sG}$ are exactly the sufficient statistics of the model (\equationref{logLinearSupervised}),
Therefore, we can optimize the supervised likelihood objective without actually
having any supervised data!
In the finite data regime, the method of moments yields the estimate
$\hat \mu^\text{mom}_\sC$ which approximates the true $\mu_\sC$.
In supervised learning, we obtain a different estimate $\hat\mu^\text{sup}_\sC$ of $\mu_\sC$ based on an empirical average
over data points.
In the limit of infinite data, both estimators converge to $\mu_\sC$.

%\algorithmref{undirected} summarizes our approach.
%In \sectionref{exclusiveViews}, we showed that $\LearnMarginals$ is a consistent
%estimator for $Z_\sC$ for any bottlenecked clique in the graphical model. 
%Consequently, \algorithmref{undirected}
%is also a consistent estimator.\footnote{Of course, the log-linear model must be identifiable.
%Identifability is violated, for example, if two components of $\phi(\vx_\sC, \vh_\sC)$ were identical.
%In general, we can identify the parameters up to the kernel of $\nabla^2 A(\theta)$.}

%\renewcommand{\algorithmicrequire}{\textbf{Input:}}
%\renewcommand{\algorithmicensure}{\textbf{Output:}}
\begin{algorithm}
  \caption{\LearnParameters}
  \label{algo:undirected}
  \begin{algorithmic}
    \REQUIRE Conditional moments $\mOpp{v}{i} = \Pr(x_v \mid h_i)$
             and hidden marginals $Z_S = \Pr(\bh_S)$.
    \ENSURE Parameters $\theta$.
    \IF{$\sG$ is directed} \STATE Normalize each $\Pr(a,\Pa(a))$. \ENDIF
    \IF{$\sG$ is undirected with low treewidth}
      \STATE Compute expected features $\{\mu_\sC\}_{\sC \in \sG}$ (\equationref{logLinearFeatures}).
      \STATE Optimize full likelihood (\equationref{logLinearSupervised}).
    \ENDIF
    \IF{$\sG$ is undirected with high treewidth}
      \STATE Compute expected features $\{\mu_S\}_{S \in \sB}$ (\equationref{logLinearFeatures}).
      \STATE Optimize pseudolikelihood (\equationref{pseudolikelihood}).
    \ENDIF
    \end{algorithmic}
\end{algorithm}

% Learning with partial moment constraints
\paragraph{Remark} If we have exclusive views for only a subset of the cliques,
we can still obtain the expected features $\mu_\sC$ for those cliques
  and use posterior regularization \citep{graca08em}, measurements \citep{liang09measurements}, or generalized
  expectation criteria \citep{mann08ge}
  to encourage $\E_{p_\theta}[\phi(\vx_\sC,\vh_\sC)]$ to match $\mu_\sC$.
The resulting objective functions would be non-convex, but we expect
  local optima to be less of an issue.
  %How to characterize this precisely is an interesting direction for future work.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Pseudolikelihood}
\label{sec:pseudolikelihood}

While we now have a complete algorithm for estimating directed and undirected
models, optimizing the full likelihood (\equationref{logLinearSupervised})
can still be computationally intractable for undirected models with high treewidth
due to the intractability of the log-partition function $A(\theta)$.
One can employ various variational approximations of $A(\theta)$ \cite{wainwright08varinf},
but these generally lead to inconsistent estimates of $\theta$.
We thus turn to an older idea of pseudolikelihood \citep{besag75pseudo}.
The pseudolikelihood objective is a sum over the log-probability of each
variable $a$ given its neighbors $\sN(a)$:
\begin{align}
  \label{eqn:pseudolikelihood}
  L_\text{pseudo}(\theta) \eqdef \E_{(\bx,\bh) \sim p_{\theta^*}}\left[\sum_{a \in X \cup H} \log p_\theta(a \mid \sN(a))\right].
\end{align}
In the fully-supervised setting, it is well-known that pseudolikelihood
provides consistent estimates which are computationally efficient
but less statistically efficiency.
(Coincidentally, this is the same high-level motivation for using method of
moments in the first place.)

Let $\phi_{a,\sN(a)}(a, \sN(a)) = \sum_{\sC \ni a} \phi_\sC(\bx_\sC,\bh_\sC)$
denote the sum over cliques $\sC$ that contain $a$; note that $\phi_{a, \sN(a)}$ only depends on $a$ and its neighbors $\sN(a)$.
We can write each conditional log-likelihood from \equationref{pseudolikelihood} as:
\begin{align*}
p_\theta(a \mid \sN(a)) = \exp(\theta^\top\phi_{a,\sN(a)}(a, \sN(a)) - A_a(\theta; \sN(a))),
\end{align*}
where the conditional log-partition function $A_a(\theta; \sN(a)) = \log \sum_{a'} \exp(\theta^\top\phi_{a,\sN(a)}(a', \sN(a)))$
involves marginalizing only over the single variable $a$.

If we knew $\mu_{a,\sN(a)} \eqdef \E[\phi_{a,\sN(a)}(a,\sN(a))]$,
then we would be able to optimize the pseudolikelihood objective again without
having access to any labeled data.
However, $\{a\}\cup\sN(a)$ does not always have exclusive views
(for example, consider $a = h_1$ and $\sN(a) = \{h_2,h_3,h_4\}$ in \figureref{examples-tree}).

However, we can decompose $\{a\}\cup\sN(a)$ as follows:
conditioning on $a$ partitions $\sN(a)$ into independent subsets;
let $\sB(a)$ be the collection of these subsets,
which we will call \emph{sub-neighborhoods}.
For example, $\sB(h_1) = \{ \{h_2\},\{h_3\},\{h_4\} \}$ in \figureref{examples-tree}
and $\sB(h_{2,2}) = \{ \{ h_{1,2}, h_{2,3}, h_{3,2}, h_{2,1} \} \}$ contains a single sub-neighborhood.

A key observation is that for each $B \in \sB(a)$, each $\{a\} \cup B$ is bidependent:
conditioning on $a$ does not introduce new independencies within $B$ by construction of $\sB(a)$,
and conditioning on any $b \in B$ does not either since every other $b' \in B \backslash \{b\}$
is connected to $a$.
Assuming $\sG$ is bottlenecked, by \lemmaref{bottleneck-views},
we have that $\{a\} \cup B$ has exclusive views.
Hence, we can recover $\Pr(a,B)$ for each $a$ and $B \in \sB(a)$.
Based on conditional independence of the sub-neighborhoods $B$ given $a$,
we have that $\Pr(a, \sN(a)) = \Pr(a) \prod_{B \in \sB(a)} \Pr(B \mid a)$.
This allows us to estimate the expected features $\mu_{a,\sN(a)}$ and use these quantities
in the optimization of the pseudolikelihood objective.

%  Therefore, we can adapt our algorithm ($\LearnMarginals$) to estimate all
%  marginals of the form $Z_{a,\sN(a)} \eqdef \Pr(a,\sN(a))$,
%  from which we can compute the features $\mu_{a,\sN(a)}$,
%  and optimize \equationref{pseudolikelihood} to obtain a consistent estimate of $\theta$.

Note that our pseudolikelihood-based approach
does depend exponentially on the size of the sub-neighborhoods,
which could be larger than the largest clique size.
Therefore,
for this approach to be effective,
each node essentially should have low degree or locally exhibits a lot of
conditional independence.
%While this method is only viable for graphs with small sub-neighboof low
%degree, i.e.\@ when $|\sN(a)|$ is small,
On the positive side, we can handle graphical models with high treewidth;
neither sample nor computational complexity depends on the treewidth.
For example, in the grid model, the treewidth is linear in the size of the grid,
but it is the degree (only 4 in this case) that governs the computational complexity.