### 2014-05-18 doc/icml2014/abstract.tex
minor edits

\begin{abstract}

Recent work on the method of moments enable
  consistent parameter estimation, but
  only for certain types of latent-variable models.
On the other hand, pure likelihood objectives, though more universally applicable,
  are difficult to optimize.
In this work, we show that using the method of moments in conjunction
  with composite likelihood
  yields consistent parameter estimates for a much broader
  class of discrete directed and undirected graphical models,
  including loopy graphs with high treewidth.
Specifically, we use tensor factorization to reveal information
about the hidden variables.
This allows us to construct convex likelihoods which can be globally optimized
to recover the parameters.
\end{abstract} 
%This renders the otherwise non-convex
%likelihood objective convex, and EM can then be used to obtain the global optimum.

%\paragraph{Last updated:} \today, \currenttime
%
%\paragraph{Errata / Incomplete}
%\begin{enumerate}
%  \item Use \texttt{\textbackslash left(}.
%  \item Use $\ell$ instead of $L$.
%  \item convex and negative log likelihood.
%\end{enumerate}
%